---
title: "environmental_data_preperation"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Preperation of GLORYS and BIOMER netCDFs obtained from [Copernicus](www.marine.copernicus.eu/services-portfolio/access-to-products/). These databases give global oceanographic condition.
UPDATE: also includes prep of the AMO and NAO data...


A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
```{r pre-install packages, message=FALSE}
packages <- c("ncdf4", "raster", "rgdal", "useful", "marmap", "sp") 
source("../src/ipak.R")
ipak(packages)
```

# NetCDFs (Biomer and GLORYS)
read in a netcdf
note this is a GLORYS dataset
```{r}
library("ncdf4")
testnc <-  "../data/env/1993_01_salinity.nc"
ncbr <- brick(testnc,lvar = 4) #lvar = 4 loads the levels
ncbr 
```

reproject ...
There is currently an issue with the projectRaster command in the raster package. Michael Sumners kindly offered a [temporary solution](https://twitter.com/mdsumner/status/1040500830539935744)
```{r}
## use ncdf4 directly
library(ncdf4)
## connection to the file
nc <- nc_open("../data/env/1993_01_salinity.nc")
## desired variable 
v <- ncvar_get(nc, "salinity")
## dimension coordinates
lon <- ncvar_get(nc, "longitude")
lat <- ncvar_get(nc, "latitude")
## a template, one map from the lon, lat, and v[,,1]
template <- raster(list(x = lon, y = lat, z = v[,,1]), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 ")
## apply template to brick, and set values to v (but flipped in Y because that's how raster works - lat is increasing in the NetCDF, so we go upside down)
diy <- setValues(brick(template), v[, ncol(v):1, ])
## target
prj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs" #note -60 used to be -91. Changed to 'straighten up'.
target <- projectExtent(diy, prj)  ## get a target extent, and we'll set the exact resolution we want
res(target) <- c(25000, 25000)
aea <- projectRaster(diy, target)
plot(aea[[1]]) #the number in brackets refers to the depth number
aea
writeRaster(aea[[1]],'../output/env/aea.tif',overwrite=TRUE,options=c('TFW=YES'))
```


if projectRaster would work properly...

#```{r}
library(raster)
canada_eq_ar_proj <- "+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
library("rgdal")
rp <- projectRaster (ncbr, crs = canada_eq_ar_proj)
rp
plot(rp[[1]])
writeRaster(rp[[1]],'../data/env/rplayer.tif',options=c('TFW=YES'), overwrite=TRUE) 
#```





----------

# create unique id layer
ok now you want to create a duplicate of the top layer and then give each cell an unique ID

First 
```{r}
aea[[1]]
unique_cell <- (aea[[1]])
```

Because the NA values are currently attributed with some other number and not NA in R (NA is set to -32767 in this particular layer), you need to make sure to exclude the cells which represent NA... so anything under 0.
the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (15,402)
```{r}
cell_val_freq <- freq(unique_cell, value=NULL)
cell_val_freq
```
ok so we have 7634 cells that have NA (or in reality a negative number that represents NA). so remove 7634 from ncell. We know that there are 15402 cells, soo... 15402-7634 = 7768 cells have data.

now give a unique ID to each cell
```{r}
unique_cell[unique_cell > 0] <- 1:7768 
writeRaster(unique_cell,'../output/env/glo_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
uniquecell_poly <- rasterToPolygons(aea[[1]], fun=NULL, n=4, na.rm=TRUE, digits=12, dissolve=FALSE)
writeOGR(uniquecell_poly, dsn = '.', layer = '../output/env/glo_unique_cell_poly', driver = "ESRI Shapefile")
```

Done!

-----------------------------------------
# The Biomer Layers...


Read in a biomer netcdf
note this only reads in the top layer which is what i need right now. To get the other layers, add ", lvar = 4" to brick (see code)
```{r}
library("ncdf4")
biobrick <- brick("../data/env/1998_01_chl.nc") # lvar = 4 
biobrick
```

Project the biomer file to match the glorys layer
```{r}
bio_aea <- projectRaster(biobrick, aea) 
bio_aea
writeRaster(bio_aea,'../output/env/bio_aea.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
```

plot
```{r}
plot(bio_aea)
```
ok now lets do the same "create unique cell ID" process for the biomer layer as you did for GLORYS. You want to check they match

Because the NA values are currently attributed with some other number and not NA in R (NA is set to -32767 in this particular layer), you need to make sure to exclude the cells which represent NA... so anything under 0.
the cell id will be a count of the cells with data. First calculate how many cells to remove from ncell (15402) 
```{r}
bio_unique_cell <- bio_aea
bio_cell_val_freq <- freq(bio_unique_cell, value=NULL)
bio_cell_val_freq
```
ok so we have 7652 cells that have NA. Note that the GLORYS layer has 7634 NA cells....  so let's remove the cells from the GLORYS raster that correspond to NA in the BIOMER raster...

```{r}
aea <- overlay(aea, bio_aea, fun = function(x, y){
  x[is.na(y[])] <- NA
  return(x)
})
plot(aea[[1]])
```


```{r}
aea_cut_cell_val_freq <- freq(aea[[1]], value=NULL)
aea_cut_cell_val_freq
```

great! So let's redo the GLORYS unique cell layer....

```{r}
unique_cell <- aea[[1]]
unique_cell[unique_cell > 0] <- 1:7750 
writeRaster(unique_cell,'../output/env/glo_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
uniquecell_poly <- rasterToPolygons(aea[[1]], fun=NULL, n=4, na.rm=TRUE, digits=12, dissolve=FALSE)
writeOGR(uniquecell_poly, dsn = '.', layer = '../output/env/glo_unique_cell_poly', driver = "ESRI Shapefile")
unique_cell
```

and create a unique cell biomer layer..

```{r}
bio_unique_cell <- bio_aea
bio_unique_cell[bio_unique_cell > -1] <- 1:7750 
writeRaster(bio_unique_cell,'../output/env/bio_unique_cell.tif',overwrite=TRUE, options=c('TFW=YES')) #so you can access it again
bio_unique_cell
```


-------------------------------------------

# AMO data 

Downloaded from [NOAA](https://www.esrl.noaa.gov/psd/data/timeseries/AMO/) (AMO, unsmoothed, short) as a txt and then converted to a .csv

load raw data
```{r}
amoraw <- read.csv("../data/env/amon.us.data.csv", header = FALSE) 
head(amoraw)
```
ok first col = year, the others represent the month. Give the columns some proper names
```{r}
colnames(amoraw) <- c("year","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
head(amoraw)
```

The data downloaded from the site indidcated that NA values are represented by -99.99. Lets replace those values with NA 
```{r}
amoraw <- replace(amoraw,amoraw==-99.99, NA)
```

Now subset the dataset to start in 1996 (the capelin data starts in 1998, but you want some previous time period stuff)
```{r}
amoraw <- amoraw[amoraw$year >= 1996, ]
head(amoraw)
```

And just for fun, plot the AMO... follow the tutorial [here](http://www.stat.pitt.edu/stoffer/tsa4/R_toot.htm)
```{r}
amocolumn <- matrix(t(amoraw[ ,2:13]), ncol=1) #month data is in col 2 - 13
amocolumn1 <- as.vector(amocolumn)
amotimeseries <- ts(amocolumn1, frequency = 12, start = c(1996,1))
plot.ts(amotimeseries)
abline(h = 0)
dev.copy(png, "../output/env/amo.png") #this prints a png of the plot
dev.off() #this turns off the print command
```


First I will create a matrix that takes all of the raw data and then appends some extra columns - year average value, year phase (warm or cold) and then seasonal values (spring, summer, autum, winter) + seasonal phase (warm cold)

Spring:	1 March - 31 May (col 4:6) Summer: 1 June - 31 August (col 7:9) Autumn: 1 Sept - 31 Nov (col 10:12) Winter: 1 Dec - 28 Feb (col 13 from prev year, 1:2)
```{r}
amophases <- amoraw 
amophases$YearAvg <- rowMeans(amophases[2:13], na.rm = TRUE) #the month data is in col 2-13
amophases$SpringAvg <-  rowMeans(amophases[4:6], na.rm = TRUE)
amophases$SummerAvg <-  rowMeans(amophases[7:9], na.rm = TRUE)
amophases$AutumnAvg <-  rowMeans(amophases[10:12], na.rm = TRUE)
```
winter has a different format - the columns are in 1 & 2, & 13 from the previous column, which means the rowmeans function needs a bit of tweaking...


1) shift the dec column you created down by 1
```{r}
library(useful)
amophases <- shift.column(data = amophases, columns = "Dec", len = 1, up = FALSE)
amophases
write.csv(amophases, file = "../output/env/amo.csv", row.names = FALSE) #for some reason you need to write to a csv and then reload for the next step to work properly...
```

the shifted data is in Dec.Shifted...
```{r}
amophases <- read.csv("../output/env/amo.csv")
amophases$WinterAvg <- rowMeans(amophases[, c("Dec.Shifted", "Jan", "Feb")], na.rm = TRUE) 
amophases
```

Now label the phases positive, negative (or avg if any happen to be at 0)
```{r}
amophases$YearPhase[amophases$YearAvg ==0] <- "avg"
amophases$YearPhase[amophases$YearAvg >0] <- "positive"
amophases$YearPhase[amophases$YearAvg <0] <- "negative"
amophases$SpringPhase[amophases$SpringAvg ==0] <- "avg"
amophases$SpringPhase[amophases$SpringAvg >0] <- "positive"
amophases$SpringPhase[amophases$SpringAvg <0] <- "negative"
amophases$SummerPhase[amophases$SummerAvg ==0] <- "avg"
amophases$SummerPhase[amophases$SummerAvg >0] <- "positive"
amophases$SummerPhase[amophases$SummerAvg <0] <- "negative"
amophases$AutumnPhase[amophases$AutumnAvg ==0] <- "avg"
amophases$AutumnPhase[amophases$AutumnAvg >0] <- "positive"
amophases$AutumnPhase[amophases$AutumnAvg <0] <- "negative"
amophases$WinterPhase[amophases$WinterAvg ==0] <- "avg"
amophases$WinterPhase[amophases$WinterAvg >0] <- "positive"
amophases$WinterPhase[amophases$WinterAvg <0] <- "negative"
```

now save this as a new csv
```{r}
write.csv(amophases, file = "../output/env/amo.csv", row.names = FALSE)
```

# NAO

Dataset mes from [NCAR](https://climatedataguide.ucar.edu/climate-data/hurrell-north-atlantic-oscillation-nao-index-pc-based). Note I am using the monthly principal component (PC)-based indices of the North Atlantic Oscillation [(NAO)](https://climatedataguide.ucar.edu/sites/default/files/nao_pc_monthly.txt). I have already converted this file into a csv and removed unncesessary text. Note there are no headers...

```{r}
naoraw <- read.csv("../data/env/amon.us.data.csv", header = FALSE) 
head(naoraw)
```

lets give some column names. col 1  year, 2:13 are the months
```{r}
colnames(naoraw) <- c("year","1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
head(naoraw)
```
 
 NA values are represented by -99.99. Change these to NA
```{r}
naoraw <- replace(naoraw,naoraw==-99.99, NA)
```
 
Now subset the dataset to start in 1996 (the capelin data starts in 1998, but you want some previous time period stuff)

```{r}
nao <- naoraw[naoraw$year >= 1996, ]
head(nao) 
```

And just for fun, plot the NAO follow the tutorial here

```{r}
naocolumn <- matrix(t(amoraw[ ,2:13]), ncol=1) #month data is in col 2 - 13
naocolumn1 <- as.vector(naocolumn)
naotimeseries <- ts(naocolumn1, frequency = 12, start = c(1996,1))
plot.ts(naotimeseries)
abline(h = 0)
dev.copy(png, "../output/env/nao.png") #this prints a png of the plot
dev.off() #this turns off the print command
```

First I will create a matrix that takes all of the raw data and then appends some extra columns - year average value, year phase (warm or cold) and then seasonal values (spring, summer, autum, winter) + seasonal phase (warm cold)

Spring:	1 March - 31 May (col 4:6) Summer: 1 June - 31 August (col 7:9) Autumn: 1 Sept - 31 Nov (col 10:12) Winter: 1 Dec - 28 Feb (col 13 from prev year, 1:2)

```{r}
naophases <- nao 
naophases$YearAvg <- rowMeans(naophases[2:13], na.rm = TRUE) #the month data is in col 2-13
naophases$SpringAvg <- rowMeans(naophases[4:6], na.rm = TRUE)
naophases$SummerAvg <- rowMeans(naophases[7:9], na.rm = TRUE)
naophases$AutumnAvg <- rowMeans(naophases[10:12], na.rm = TRUE)
```

winter has a different format - the columns are in 1 & 2, & 13 from the previous column, which means the rowmeans function needs a bit of tweaking...

shift the dec column you created down by 1
```{r}
library(useful)
naophases <- shift.column(data = naophases, columns = "12", len = 1, up = FALSE)
naophases
write.csv(naophases, file = "../output/env/nao.csv", row.names = FALSE) #for some reason you need to write to a csv and then reload for the next step to work properly...
```

he shifted data is in 12.Shifted...
```{r}
naophases <- read.csv("../output/env/nao.csv")
naophases$WinterAvg <- rowMeans(naophases[ , c("X12.Shifted", "X1", "X2")], na.rm = TRUE)  #x because apparently all the col names start with x for some reason...
naophases
```

Now label the phases positive, negative (or avg if any happen to be at 0)

```{r}
naophases$YearPhase[naophases$YearAvg ==0] <- "avg"
naophases$YearPhase[naophases$YearAvg >0] <- "positive"
naophases$YearPhase[naophases$YearAvg <0] <- "negative"
naophases$SpringPhase[naophases$SpringAvg ==0] <- "avg"
naophases$SpringPhase[naophases$SpringAvg >0] <- "positive"
naophases$SpringPhase[naophases$SpringAvg <0] <- "negative"
naophases$SummerPhase[naophases$SummerAvg ==0] <- "avg"
naophases$SummerPhase[naophases$SummerAvg >0] <- "positive"
naophases$SummerPhase[naophases$SummerAvg <0] <- "negative"
naophases$AutumnPhase[naophases$AutumnAvg ==0] <- "avg"
naophases$AutumnPhase[naophases$AutumnAvg >0] <- "positive"
naophases$AutumnPhase[naophases$AutumnAvg <0] <- "negative"
naophases$WinterPhase[naophases$WinterAvg ==0] <- "avg"
naophases$WinterPhase[naophases$WinterAvg >0] <- "positive"
naophases$WinterPhase[naophases$WinterAvg <0] <- "negative"
```

now save this as a new csv

```{r}
write.csv(naophases, file = "../output/env/nao.csv", row.names = FALSE)
```

# Bathymetric

bottom depth from [GEBCO](www.gebco.net)

```{r}
library(ncdf4)
library(marmap)
bat <-readGEBCO.bathy("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
summary(bat)
```

create pretty custom color palette and then plot the bathy laye
```{r}
blues <- colorRampPalette(c("lightblue", "cadetblue2", "cadetblue1", "white"))
plot(bat, n = 1, image = TRUE, bpal = blues(100))
```
lovely jubbely

add the species data...
```{r}
data_aea <-read.csv( "../output/bio/data_aea_cell_amo_nao_bins_env.csv", header = TRUE) #just for canada pc as not loaded
plot(bat, n = 1, image = TRUE, bpal = blues(100))
points(data_aea$decimalLongitude, data_aea$decimalLatitude, pch = 20, col = "yellow") #note these are both the wgs84 images
```
and now get depth... note all the points and bathy data is wgs!

```{r}
data_aea$bottom_depth <- get.depth(bat, data_aea[ ,5:4], locator = FALSE)
data_aea$bottom_depth <- data_aea$bottom_depth[, 3] #this step is because get.bath actually produces it's own dataframe with long, lat, and depth. This is extracting the depth column and deplacing data_aea$bottom_depth with the correct data.
head(data_aea)
```

```{r}
write.csv(data_aea,"../output/bio/data_aea_cell_amo_nao_bins_btmdepth.csv", row.names = FALSE)
```

And now create a raster of Gebco in aea for mapping once the SDM is done...

```{r}
library(raster)
gebco <- raster("../data/env/GEBCO_2014_2D_-70.0_34.0_-43.0_70.0.nc")
gebco
gebco_aea <- projectRaster(gebco, aea)  
gebco_aea
```

```{r}
plot(gebco_aea)
```

# attach environmental correlates

This loop runs through the netcdf files and then looks for which rows in data_aea it should extract the value to point from, and at what depth (netcDF layer)

First you need to put the observations (data_aea) into a spatialpointsdataframe.
```{r}
library(sp)
xy <- data_aea[ ,c(18,19)] # This is to tell R where the coordinates are (in column 18 and 19). Note that the column order needs to be longitude, latitude
data_aea_sp <- SpatialPointsDataFrame(coords = xy, data = data_aea, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
```

```{r}
netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not includedCores
library (ncdf4)
library(sp)
library(rgdal)
library(raster)
for (i in 1:no_netcdf) {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(data_aea_sp)) {  
      print (j)
      de <- data_aea_sp$depth[[j]]  # a variable for the observation depth layer
      yr <- (data_aea_sp$year[j])  # a variable for the observation year
      mth <- (data_aea_sp$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
           
             data_aea_sp$temp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
              if (is.na(de)){
                data_aea_sp$temp_depth[j] <- NA
              } else  
                data_aea_sp$temp_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
              data_aea_sp$salinity_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
              if (is.na(de)){
                data_aea_sp$salinity_depth[j] <- NA
              } else  
                data_aea_sp$salinity_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
              data_aea_sp$chl_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
              if (is.na(de)){
                data_aea_sp$chl_depth[j] <- NA
              } else  
                data_aea_sp$chl_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
              data_aea_sp$o2_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
              if (is.na(de)){
                data_aea_sp$o2_depth[j] <- NA
              } else  
                data_aea_sp$o2_depth[j] <- extract(x=temp_brick[[de]], y = data_aea_sp[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
              data_aea_sp$mlp_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
              data_aea_sp$ssh_surface[j] <- extract(x=temp_brick[[1]], y = data_aea_sp[j, ]) 
            
          }
     
    }
}
```