---
title: "merging_datasets"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
merging background and presence data into one giant dataset, and timesliced subsets


A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there are probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
```{r pre-install packages, message=FALSE}
packages <- c("plyr", "xlsx" , "dplyr") 
source("../src/ipak.R")
ipak(packages)
```

read in the presence and background
```{r read in presence and background}
presencemerging <- read.csv("../output/bio/presence_points_without_envdata_relooped_glbathy_nodup.csv", header = TRUE)
backgroundmerging <- read.csv("../output/bio/background_complete_obs_cels_globot_cellid_xyzt.csv", header = TRUE)
```

now check 
```{r}
colnames(presencemerging)
```

```{r}
colnames(backgroundmerging)
```

ok so the two dataframed need a spot of cleaning

```{r removing crap columns from presence and background}
presencemerging <- subset(presencemerging, select = -c(longitude_meters.1, latitude_meters.1, longitude_meters.2, latitude_meters.2, XXtotal_cell_obs_xyzt))
backgroundmerging <- subset(backgroundmerging, select = -c(longitude_meters.1, latitude_meters.1, longitude_meters.2, latitude_meters.2, X, optional, optional.1))
```

for consistency, rename a couple of presencemerging columns
```{r renaming presence columns}
names(presencemerging)[names(presencemerging)=="total_cell_obs"] <- "total_cell_obs_xy"
names(presencemerging)[names(presencemerging)=="yymm_cell_obs"] <- "total_cell_obs_xyt"
```

now merge the two datasets
```{r merge pres and background}
presab <- rbind.fill(presencemerging, backgroundmerging)
write.csv(presab, "../output/bio/presab.csv", row.names = FALSE)
head(presab)
```

lovely jubley


# weighted 10k for each month

For each month, the 10,000 background points will be garnered from background points created for each month-year and weighted to the observations. For example, for January 22% of the unique observations by cell occurred in 2003, 49% in 2004, 7% in 2005, and 21% in 2006. Thus, 22% of the background points for the January model will come from 2003, 49% from 2004, 7% in 2005, and 21% in 2006. 
See ‘Percentage of occurrences by unique cell (xyzt)’ in Appendix Observations per Cell in results document + excel + data_exploration.

Some other analysis has indicated that background points have been gathered from 1BCDEF (ouside Canada EEZ and inside Greenland EEZ). Just do a table here

```{r}
nafounique <- unique(backgroundmerging$nafo_zone)
nafounique
nafo1b <- subset(backgroundmerging, nafo_zone == "1B")
nafo1bc <- subset(backgroundmerging, nafo_zone == "1C")
nafo1d <- subset(backgroundmerging, nafo_zone == "1D")
nafo1be <- subset(backgroundmerging, nafo_zone == "1E")
nafo1f <- subset(backgroundmerging, nafo_zone == "1F")
nafo1 <- rbind(nafo1b, nafo1bc, nafo1be, nafo1f)
nrow(nafo1) #140174
```

and the spread of these funny nafo points in year month...

```{r background nafo1x by year-month}
nafo1_yymm <- with(nafo1, table(year, month))
write.csv(nafo1_yymm, file = "../output/bio/background_nafo1x_mth_yr.csv")
nafo1_yymm
```

ok so nothing too crazy... Orignially i removed these points, but I now think they should be left in to honour the MCP (though it is really shapped by two points). as per slr email

code was... 
backno1 <- backgroundmerging[!(backgroundmerging$nafo_zone == "1B"),]
backno1 <- backno1[!(backno1$nafo_zone == "1C"),]
backno1 <- backno1[!(backno1$nafo_zone == "1D"),]
backno1 <- backno1[!(backno1$nafo_zone == "1E"),]
backno1 <- backno1[!(backno1$nafo_zone == "1F"),]

nafounique <- unique(backno1$nafo_zone)
nafounique
head(backno1)


ok now randomly select background points from each timeslice as per this table (see results doc appendix observations per cell + excel spreadsheet)

	    1	2	3	4	5	6	7	8	9	10	11	12
1998	0	0	0	846	790	835	518	0	358	1315	1123	107
1999	0	0	1433	603	595	968	648	0	620	169	835	551
2000	0	0	1277	603	537	746	492	0	652	590	1025	725
2001	0	5000	748	990	751	637	389	0	390	405	495	1430
2002	0	0	996	1062	761	676	440	0	811	624	420	966
2003	2222	0	717	1004	888	809	492	0	318	590	311	1014
2004	4938	0	0	732	790	797	674	1004	652	472	604	1101
2005	700	4000	872	316	644	676	1010	1270	556	776	858	416
2006	2140	1000	1558	301	49	790	674	1410	588	1096	616	783
2007	0	0	623	746	508	593	1813	1404	636	624	553	821
2008	0	0	934	516	741	797	777	1418	564	455	748	589
2009	0	0	312	946	1405	351	207	582	580	556	915	454
2010	0	0	530	560	663	720	648	596	747	1096	875	560
2011	0	0	0	775	878	605	285	758	533	1062	593	483
2012	0	0	0	0	0	0	208	793	604	0	0	0
2013	0	0	0	0	0	0	699	660	509	119	12	0
2014	0	0	0	0	0	0	26	77	818	51	0	0
2015	0	0	0	0	0	0	0	28	64	0	17	0
Total	10000	10000	10000	10000	10000	10000	10000	10000	10000	10000	10000	10000

Ok so for Jan:
2003: 2222
2004: 4938
2005: 700
2006: 2140

```{r}
jan2003 <- subset(backno1, year == "2003" & month == "1")
jan2003 <- sample_n(jan2003, 2222) 
jan2004 <- subset(backno1, year == "2004" & month == "1")
jan2004 <- sample_n(jan2004, 4938) 
jan2005 <- subset(backno1, year == "2005" & month == "1")
jan2005 <- sample_n(jan2005, 700) 
jan2006 <- subset(backno1, year == "2006" & month == "1")
jan2006 <- sample_n(jan2006, 2140) 
jan <- rbind(jan2003, jan2004, jan2005, jan2006)
```

for feb
2001: 5000
2005: 4000
2006: 1000

```{r}
feb2001 <- subset(backno1, year == "2001" & month == "2")
feb2001 <- sample_n(feb2001, 5000) 
feb2005 <- subset(backno1, year == "2005" & month == "2")
feb2005 <- sample_n(feb2005, 4000) 
feb2006 <- subset(backno1, year == "2006" & month == "2")
feb2006 <- sample_n(feb2006, 1000) 
feb <- rbind(feb2001, feb2005, feb2006)
```

for march
1999	1433
2000	1277
2001	748
2002	996
2003	717
2005	872
2006	1558
2007	623
2008	934
2009	312
2010	530

```{r}
mar1999 <- subset(backno1, year == "1999" & month == "3")
mar1999 <- sample_n(mar1999, 1433) 
mar2000 <- subset(backno1, year == "2000" & month == "3")
mar2000 <- sample_n(mar2000, 1277) 
mar2001 <- subset(backno1, year == "2001" & month == "3")
mar2001 <- sample_n(mar2001, 748)
mar2002 <- subset(backno1, year == "2002" & month == "3")
mar2002 <- sample_n(mar2002, 996)
mar2003 <- subset(backno1, year == "2003" & month == "3")
mar2003 <- sample_n(mar2003, 717)
mar2005 <- subset(backno1, year == "2005" & month == "3")
mar2005 <- sample_n(mar2005, 872)
mar2006 <- subset(backno1, year == "2006" & month == "3")
mar2006 <- sample_n(mar2006, 1558)
mar2007 <- subset(backno1, year == "2007" & month == "3")
mar2007 <- sample_n(mar2007, 623)
mar2008 <- subset(backno1, year == "2008" & month == "3")
mar2008 <- sample_n(mar2008, 934)
mar2009 <- subset(backno1, year == "2009" & month == "3")
mar2009 <- sample_n(mar2009, 312)
mar2010 <- subset(backno1, year == "2010" & month == "3")
mar2010 <- sample_n(mar2010, 530)
mar <- rbind(mar1999, mar2000, mar2001, mar2002, mar2003, mar2005, mar2006, mar2007, mar2008, mar2009, mar2010)
```

For april
1998	846
1999	603
2000	603
2001	990
2002	1062
2003	1004
2004	732
2005	316
2006	301
2007	746
2008	516
2009	946
2010	560
2011	775

```{r}
apr1998 <- subset(backno1, year == "1998" & month == "4")
apr1998 <- sample_n(apr1998, 846) 
apr1999 <- subset(backno1, year == "1999" & month == "4")
apr1999 <- sample_n(apr1999, 603) 
apr2000 <- subset(backno1, year == "2000" & month == "4")
apr2000 <- sample_n(apr2000, 603) 
apr2001 <- subset(backno1, year == "2001" & month == "4")
apr2001 <- sample_n(apr2001, 990)
apr2002 <- subset(backno1, year == "2002" & month == "4")
apr2002 <- sample_n(apr2002, 1062)
apr2003 <- subset(backno1, year == "2003" & month == "4")
apr2003 <- sample_n(apr2003, 1004)
apr2004 <- subset(backno1, year == "2004" & month == "4")
apr2004 <- sample_n(apr2004, 732)
apr2005 <- subset(backno1, year == "2005" & month == "4")
apr2005 <- sample_n(apr2005, 316)
apr2006 <- subset(backno1, year == "2006" & month == "4")
apr2006 <- sample_n(apr2006, 301)
apr2007 <- subset(backno1, year == "2007" & month == "4")
apr2007 <- sample_n(apr2007, 746)
apr2008 <- subset(backno1, year == "2008" & month == "4")
apr2008 <- sample_n(apr2008, 516)
apr2009 <- subset(backno1, year == "2009" & month == "4")
apr2009 <- sample_n(apr2009, 946)
apr2010 <- subset(backno1, year == "2010" & month == "4")
apr2010 <- sample_n(apr2010, 560)
apr2011 <- subset(backno1, year == "2011" & month == "4")
apr2011 <- sample_n(apr2011, 775)
apr <- rbind(apr1998, apr1999, apr2000, apr2001, apr2002, apr2003, apr2004, apr2005, apr2006, apr2007, apr2008, apr2009, apr2010, apr2011)
```


For May
1998	790
1999	595
2000	537
2001	751
2002	761
2003	888
2004	790
2005	644
2006	49
2007	508
2008	741
2009	1405
2010	663
2011	878

```{r}
may1998 <- subset(backno1, year == "1998" & month == "5")
may1998 <- sample_n(may1998, 790) 
may1999 <- subset(backno1, year == "1999" & month == "5")
may1999 <- sample_n(may1999, 595) 
may2000 <- subset(backno1, year == "2000" & month == "5")
may2000 <- sample_n(may2000, 537) 
may2001 <- subset(backno1, year == "2001" & month == "5")
may2001 <- sample_n(may2001, 751)
may2002 <- subset(backno1, year == "2002" & month == "5")
may2002 <- sample_n(may2002, 761)
may2003 <- subset(backno1, year == "2003" & month == "5")
may2003 <- sample_n(may2003, 888)
may2004 <- subset(backno1, year == "2004" & month == "5")
may2004 <- sample_n(may2004, 790)
may2005 <- subset(backno1, year == "2005" & month == "5")
may2005 <- sample_n(may2005, 644)
may2006 <- subset(backno1, year == "2006" & month == "5")
may2006 <- sample_n(may2006, 49)
may2007 <- subset(backno1, year == "2007" & month == "5")
may2007 <- sample_n(may2007, 508)
may2008 <- subset(backno1, year == "2008" & month == "5")
may2008 <- sample_n(may2008, 741)
may2009 <- subset(backno1, year == "2009" & month == "5")
may2009 <- sample_n(may2009, 1405)
may2010 <- subset(backno1, year == "2010" & month == "5")
may2010 <- sample_n(may2010, 663)
may2011 <- subset(backno1, year == "2011" & month == "5")
may2011 <- sample_n(may2011, 878)
may <- rbind(may1998, may1999, may2000, may2001, may2002, may2003, may2004, may2005, may2006, may2007, may2008, may2009, may2010, may2011)
```

for June

1998	835
1999	968
2000	746
2001	637
2002	676
2003	809
2004	797
2005	676
2006	790
2007	593
2008	797
2009	351
2010	720
2011	605

```{r}
jun1998 <- subset(backno1, year == "1998" & month == "6")
jun1998 <- sample_n(jun1998, 835) 
jun1999 <- subset(backno1, year == "1999" & month == "6")
jun1999 <- sample_n(jun1999, 968) 
jun2000 <- subset(backno1, year == "2000" & month == "6")
jun2000 <- sample_n(jun2000, 746) 
jun2001 <- subset(backno1, year == "2001" & month == "6")
jun2001 <- sample_n(jun2001, 637)
jun2002 <- subset(backno1, year == "2002" & month == "6")
jun2002 <- sample_n(jun2002, 676)
jun2003 <- subset(backno1, year == "2003" & month == "6")
jun2003 <- sample_n(jun2003, 809)
jun2004 <- subset(backno1, year == "2004" & month == "6")
jun2004 <- sample_n(jun2004, 797)
jun2005 <- subset(backno1, year == "2005" & month == "6")
jun2005 <- sample_n(jun2005, 676)
jun2006 <- subset(backno1, year == "2006" & month == "6")
jun2006 <- sample_n(jun2006, 790)
jun2007 <- subset(backno1, year == "2007" & month == "6")
jun2007 <- sample_n(jun2007, 593)
jun2008 <- subset(backno1, year == "2008" & month == "6")
jun2008 <- sample_n(jun2008, 797)
jun2009 <- subset(backno1, year == "2009" & month == "6")
jun2009 <- sample_n(jun2009, 351)
jun2010 <- subset(backno1, year == "2010" & month == "6")
jun2010 <- sample_n(jun2010, 720)
jun2011 <- subset(backno1, year == "2011" & month == "6")
jun2011 <- sample_n(jun2011, 605)
jun <- rbind(jun1998, jun1999, jun2000, jun2001, jun2002, jun2003, jun2004, jun2005, jun2006, jun2007, jun2008, jun2009, jun2010, jun2011)
```

July

1998	518
1999	648
2000	492
2001	389
2002	440
2003	492
2004	674
2005	1010
2006	674
2007	1813
2008	777
2009	207
2010	648
2011	285
2012	208
2013	699
2014	26

```{r}
jul1998 <- subset(backno1, year == "1998" & month == "7")
jul1998 <- sample_n(jul1998, 518) 
jul1999 <- subset(backno1, year == "1999" & month == "7")
jul1999 <- sample_n(jul1999, 648) 
jul2000 <- subset(backno1, year == "2000" & month == "7")
jul2000 <- sample_n(jul2000, 492) 
jul2001 <- subset(backno1, year == "2001" & month == "7")
jul2001 <- sample_n(jul2001, 389)
jul2002 <- subset(backno1, year == "2002" & month == "7")
jul2002 <- sample_n(jul2002, 440)
jul2003 <- subset(backno1, year == "2003" & month == "7")
jul2003 <- sample_n(jul2003, 492)
jul2004 <- subset(backno1, year == "2004" & month == "7")
jul2004 <- sample_n(jul2004, 674)
jul2005 <- subset(backno1, year == "2005" & month == "7")
jul2005 <- sample_n(jul2005, 1010)
jul2006 <- subset(backno1, year == "2006" & month == "7")
jul2006 <- sample_n(jul2006, 674)
jul2007 <- subset(backno1, year == "2007" & month == "7")
jul2007 <- sample_n(jul2007, 1813)
jul2008 <- subset(backno1, year == "2008" & month == "7")
jul2008 <- sample_n(jul2008, 777)
jul2009 <- subset(backno1, year == "2009" & month == "7")
jul2009 <- sample_n(jul2009, 207)
jul2010 <- subset(backno1, year == "2010" & month == "7")
jul2010 <- sample_n(jul2010, 648)
jul2011 <- subset(backno1, year == "2011" & month == "7")
jul2011 <- sample_n(jul2011, 285)
jul2012 <- subset(backno1, year == "2012" & month == "7")
jul2012 <- sample_n(jul2012, 208)
jul2013 <- subset(backno1, year == "2013" & month == "7")
jul2013 <- sample_n(jul2013, 699)
jul2014 <- subset(backno1, year == "2014" & month == "7")
jul2014 <- sample_n(jul2014, 26)
jul <- rbind(jul1998, jul1999, jul2000, jul2001, jul2002, jul2003, jul2004, jul2005, jul2006, jul2007, jul2008, jul2009, jul2010, jul2011, jul2012, jul2013, jul2014)
```

For Aug
2004	1004
2005	1270
2006	1410
2007	1404
2008	1418
2009	582
2010	596
2011	758
2012	793
2013	660
2014	77
2015	28


```{r}
aug2004 <- subset(backno1, year == "2004" & month == "8")
aug2004 <- sample_n(aug2004, 1004)
aug2005 <- subset(backno1, year == "2005" & month == "8")
aug2005 <- sample_n(aug2005, 1270)
aug2006 <- subset(backno1, year == "2006" & month == "8")
aug2006 <- sample_n(aug2006, 1410)
aug2007 <- subset(backno1, year == "2007" & month == "8")
aug2007 <- sample_n(aug2007, 1404)
aug2008 <- subset(backno1, year == "2008" & month == "8")
aug2008 <- sample_n(aug2008, 1418)
aug2009 <- subset(backno1, year == "2009" & month == "8")
aug2009 <- sample_n(aug2009, 582)
aug2010 <- subset(backno1, year == "2010" & month == "8")
aug2010 <- sample_n(aug2010, 596)
aug2011 <- subset(backno1, year == "2011" & month == "8")
aug2011 <- sample_n(aug2011, 758)
aug2012 <- subset(backno1, year == "2012" & month == "8")
aug2012 <- sample_n(aug2012, 793)
aug2013 <- subset(backno1, year == "2013" & month == "8")
aug2013 <- sample_n(aug2013, 660)
aug2014 <- subset(backno1, year == "2014" & month == "8")
aug2014 <- sample_n(aug2014, 77)
aug2015 <- subset(backno1, year == "2015" & month == "8")
aug2015 <- sample_n(aug2015, 28)
aug <- rbind(aug2004, aug2005, aug2006, aug2007, aug2008, aug2009, aug2010, aug2011, aug2012, aug2013, aug2014, aug2015)
```

For sep
1998	358
1999	620
2000	652
2001	390
2002	811
2003	318
2004	652
2005	556
2006	588
2007	636
2008	564
2009	580
2010	747
2011	533
2012	604
2013	509
2014	818
2015	64

```{r}
sep1998 <- subset(backno1, year == "1998" & month == "9")
sep1998 <- sample_n(sep1998, 358) 
sep1999 <- subset(backno1, year == "1999" & month == "9")
sep1999 <- sample_n(sep1999, 620) 
sep2000 <- subset(backno1, year == "2000" & month == "9")
sep2000 <- sample_n(sep2000, 652) 
sep2001 <- subset(backno1, year == "2001" & month == "9")
sep2001 <- sample_n(sep2001, 390)
sep2002 <- subset(backno1, year == "2002" & month == "9")
sep2002 <- sample_n(sep2002, 811)
sep2003 <- subset(backno1, year == "2003" & month == "9")
sep2003 <- sample_n(sep2003, 318)
sep2004 <- subset(backno1, year == "2004" & month == "9")
sep2004 <- sample_n(sep2004, 652)
sep2005 <- subset(backno1, year == "2005" & month == "9")
sep2005 <- sample_n(sep2005, 556)
sep2006 <- subset(backno1, year == "2006" & month == "9")
sep2006 <- sample_n(sep2006, 588)
sep2007 <- subset(backno1, year == "2007" & month == "9")
sep2007 <- sample_n(sep2007, 636)
sep2008 <- subset(backno1, year == "2008" & month == "9")
sep2008 <- sample_n(sep2008, 564)
sep2009 <- subset(backno1, year == "2009" & month == "9")
sep2009 <- sample_n(sep2009, 580)
sep2010 <- subset(backno1, year == "2010" & month == "9")
sep2010 <- sample_n(sep2010, 747)
sep2011 <- subset(backno1, year == "2011" & month == "9")
sep2011 <- sample_n(sep2011, 533)
sep2012 <- subset(backno1, year == "2012" & month == "9")
sep2012 <- sample_n(sep2012, 604)
sep2013 <- subset(backno1, year == "2013" & month == "9")
sep2013 <- sample_n(sep2013, 509)
sep2014 <- subset(backno1, year == "2014" & month == "9")
sep2014 <- sample_n(sep2014, 818)
sep2015 <- subset(backno1, year == "2015" & month == "9")
sep2015 <- sample_n(sep2015, 64)
sep <- rbind(sep1998, sep1999, sep2000, sep2001, sep2002, sep2003, sep2004, sep2005, sep2006, sep2007, sep2008, sep2009, sep2010, sep2011, sep2012, sep2013, sep2014, sep2015)
```


For oct
1998	1315
1999	169
2000	590
2001	405
2002	624
2003	590
2004	472
2005	776
2006	1096
2007	624
2008	455
2009	556
2010	1096
2011	1062
2013	119
2014	51


```{r}
oct1998 <- subset(backno1, year == "1998" & month == "10")
oct1998 <- sample_n(oct1998, 1315) 
oct1999 <- subset(backno1, year == "1999" & month == "10")
oct1999 <- sample_n(oct1999, 169) 
oct2000 <- subset(backno1, year == "2000" & month == "10")
oct2000 <- sample_n(oct2000, 590) 
oct2001 <- subset(backno1, year == "2001" & month == "10")
oct2001 <- sample_n(oct2001, 405)
oct2002 <- subset(backno1, year == "2002" & month == "10")
oct2002 <- sample_n(oct2002, 624)
oct2003 <- subset(backno1, year == "2003" & month == "10")
oct2003 <- sample_n(oct2003, 590)
oct2004 <- subset(backno1, year == "2004" & month == "10")
oct2004 <- sample_n(oct2004, 472)
oct2005 <- subset(backno1, year == "2005" & month == "10")
oct2005 <- sample_n(oct2005, 776)
oct2006 <- subset(backno1, year == "2006" & month == "10")
oct2006 <- sample_n(oct2006, 1096)
oct2007 <- subset(backno1, year == "2007" & month == "10")
oct2007 <- sample_n(oct2007, 624)
oct2008 <- subset(backno1, year == "2008" & month == "10")
oct2008 <- sample_n(oct2008, 455)
oct2009 <- subset(backno1, year == "2009" & month == "10")
oct2009 <- sample_n(oct2009, 556)
oct2010 <- subset(backno1, year == "2010" & month == "10")
oct2010 <- sample_n(oct2010, 1096)
oct2011 <- subset(backno1, year == "2011" & month == "10")
oct2011 <- sample_n(oct2011, 1062)
oct2013 <- subset(backno1, year == "2013" & month == "10")
oct2013 <- sample_n(oct2013, 119)
oct2014 <- subset(backno1, year == "2014" & month == "10")
oct2014 <- sample_n(oct2014, 51)
oct <- rbind(oct1998, oct1999, oct2000, oct2001, oct2002, oct2003, oct2004, oct2005, oct2006, oct2007, oct2008, oct2009, oct2010, oct2011, oct2013, oct2014)
```

Nov
1998	1123
1999	835
2000	1025
2001	495
2002	420
2003	311
2004	604
2005	858
2006	616
2007	553
2008	748
2009	915
2010	875
2011	593
2013	12
2015	17

```{r}
nov1998 <- subset(backno1, year == "1998" & month == "11")
nov1998 <- sample_n(nov1998, 1123) 
nov1999 <- subset(backno1, year == "1999" & month == "11")
nov1999 <- sample_n(nov1999, 835) 
nov2000 <- subset(backno1, year == "2000" & month == "11")
nov2000 <- sample_n(nov2000, 1025) 
nov2001 <- subset(backno1, year == "2001" & month == "11")
nov2001 <- sample_n(nov2001, 495)
nov2002 <- subset(backno1, year == "2002" & month == "11")
nov2002 <- sample_n(nov2002, 420)
nov2003 <- subset(backno1, year == "2003" & month == "11")
nov2003 <- sample_n(nov2003, 311)
nov2004 <- subset(backno1, year == "2004" & month == "11")
nov2004 <- sample_n(nov2004, 604)
nov2005 <- subset(backno1, year == "2005" & month == "11")
nov2005 <- sample_n(nov2005, 858)
nov2006 <- subset(backno1, year == "2006" & month == "11")
nov2006 <- sample_n(nov2006, 616)
nov2007 <- subset(backno1, year == "2007" & month == "11")
nov2007 <- sample_n(nov2007, 553)
nov2008 <- subset(backno1, year == "2008" & month == "11")
nov2008 <- sample_n(nov2008, 748)
nov2009 <- subset(backno1, year == "2009" & month == "11")
nov2009 <- sample_n(nov2009, 915)
nov2010 <- subset(backno1, year == "2010" & month == "11")
nov2010 <- sample_n(nov2010, 875)
nov2011 <- subset(backno1, year == "2011" & month == "11")
nov2011 <- sample_n(nov2011, 593)
nov2013 <- subset(backno1, year == "2013" & month == "11")
nov2013 <- sample_n(nov2013, 12)
nov2015 <- subset(backno1, year == "2015" & month == "11")
nov2015 <- sample_n(nov2015, 17)
nov <- rbind(nov1998, nov1999, nov2000, nov2001, nov2002, nov2003, nov2004, nov2005, nov2006, nov2007, nov2008, nov2009, nov2010, nov2011, nov2013, nov2015)
```

Dec
1998	107
1999	551
2000	725
2001	1430
2002	966
2003	1014
2004	1101
2005	416
2006	783
2007	821
2008	589
2009	454
2010	560
2011	483


```{r}
dec1998 <- subset(backno1, year == "1998" & month == "12")
dec1998 <- sample_n(dec1998, 107) 
dec1999 <- subset(backno1, year == "1999" & month == "12")
dec1999 <- sample_n(dec1999, 551) 
dec2000 <- subset(backno1, year == "2000" & month == "12")
dec2000 <- sample_n(dec2000, 725) 
dec2001 <- subset(backno1, year == "2001" & month == "12")
dec2001 <- sample_n(dec2001, 1430)
dec2002 <- subset(backno1, year == "2002" & month == "12")
dec2002 <- sample_n(dec2002, 966)
dec2003 <- subset(backno1, year == "2003" & month == "12")
dec2003 <- sample_n(dec2003, 1014)
dec2004 <- subset(backno1, year == "2004" & month == "12")
dec2004 <- sample_n(dec2004, 1101)
dec2005 <- subset(backno1, year == "2005" & month == "12")
dec2005 <- sample_n(dec2005, 416)
dec2006 <- subset(backno1, year == "2006" & month == "12")
dec2006 <- sample_n(dec2006, 783)
dec2007 <- subset(backno1, year == "2007" & month == "12")
dec2007 <- sample_n(dec2007, 821)
dec2008 <- subset(backno1, year == "2008" & month == "12")
dec2008 <- sample_n(dec2008, 589)
dec2009 <- subset(backno1, year == "2009" & month == "12")
dec2009 <- sample_n(dec2009, 454)
dec2010 <- subset(backno1, year == "2010" & month == "12")
dec2010 <- sample_n(dec2010, 560)
dec2011 <- subset(backno1, year == "2011" & month == "12")
dec2011 <- sample_n(dec2011, 483)
dec <- rbind(dec1998, dec1999, dec2000, dec2001, dec2002, dec2003, dec2004, dec2005, dec2006, dec2007, dec2008, dec2009, dec2010, dec2011)
```


#select points weighted to year and nafo

as per discussion with MJ, modify the backgroudn point selection. file with numbers: background_point_dist_yr_nafo - based on # points per nafo region

```{r}
backno1 <- read.csv("../output/bio/presab.csv", header = TRUE)
backno1 <- subset(backno1, occurrence == "0")
head(backno1)
```


Ok so for Jan 2003: 2222 points from:

0A	52
0B	215
1B	4
1C	19
1D	34
1E	45
1F	85
2G	151
2H	160
2J	189
3K	196
3L	165
3M	10
3N	60
3O	86
3Pn	8
3Ps	89
4R	43
4S	81
4T	113
4Vn	24
4Vs	89
4W	90
4X	84
5Y	33
5Ze	11
HudsonStrait	89

```{r jan 2003}
jan20030a <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "0A")
jan20030a <- sample_n(jan20030a, 51) 
jan20030b <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "0B")
jan20030b <- sample_n(jan20030b, 215) 
jan20031B <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "1B")
jan20031B <- sample_n(jan20031B, 4)
jan20031C <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "1C")
jan20031C <- sample_n(jan20031C, 19) 
jan20031d <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "1D")
jan20031d <- sample_n(jan20031d, 34)
jan20031E <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "1E")
jan20031E <- sample_n(jan20031E, 45)
jan20031F <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "1F")
jan20031F <- sample_n(jan20031F, 85)
jan20032G <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "2G")
jan20032G <- sample_n(jan20032G, 151) 
jan20032H <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "2H")
jan20032H <- sample_n(jan20032H, 160) 
jan20032J <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "2J")
jan20032J <- sample_n(jan20032J, 189) 
jan20033k <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "3K")
jan20033k <- sample_n(jan20033k, 196) 
jan20033L <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "3L")
jan20033L <- sample_n(jan20033L, 164)
jan20033M <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "3M")
jan20033M <- sample_n(jan20033M, 10)
jan20033N <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "3N")
jan20033N <- sample_n(jan20033N, 60) 
jan20033o <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "3O")
jan20033o <- sample_n(jan20033o, 86) 
jan20033pn <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "3Pn")
jan20033pn <- sample_n(jan20033pn, 8) 
jan20033ps <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "3Ps")
jan20033ps <- sample_n(jan20033ps, 89) 
jan20034r <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "4R")
jan20034r <- sample_n(jan20034r, 43) 
jan20034s <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "4S")
jan20034s <- sample_n(jan20034s, 81) 
jan20034t <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "4T")
jan20034t <- sample_n(jan20034t, 113) 
jan20034vn <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "4Vn")
jan20034vn <- sample_n(jan20034vn, 23) 
jan20034vs <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "4Vs")
jan20034vs <- sample_n(jan20034vs, 89) 
jan20034w <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "4W")
jan20034w <- sample_n(jan20034w, 90) 
jan20034x <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "4X")
jan20034x <- sample_n(jan20034x, 84) 
jan20035y <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "5Y")
jan20035y <- sample_n(jan20035y, 33) 
jan20035ze <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "5Ze")
jan20035ze <- sample_n(jan20035ze, 11) 
jan2003hud <- subset(backno1, year == "2003" & month == "1" & nafo_zone == "HudsonStrait")
jan2003hud <- sample_n(jan2003hud, 89) 
jan2003 <- rbind(jan20030a, jan20030b, jan20031B, jan20031C, jan20031d, jan20031E, jan20031F,  jan20032G, jan20032H, jan20032J, jan20033k, jan20033L, jan20033M, jan20033N, jan20033o, jan20033pn, jan20033ps, jan20034r, jan20034s, jan20034t, jan20034vn, jan20034vs, jan20034w, jan20034x, jan20035y, jan20035ze, jan2003hud)
```

```{r jan 2004}
jan20040a <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "0A")
jan20040a <- sample_n(jan20040a, 117) 
jan20040b <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "0B")
jan20040b <- sample_n(jan20040b, 478) 
jan20041B <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "1B")
jan20041B <- sample_n(jan20041B, 9)
jan20041C <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "1C")
jan20041C <- sample_n(jan20041C, 41) 
jan20041d <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "1D")
jan20041d <- sample_n(jan20041d, 77)
jan20041E <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "1E")
jan20041E <- sample_n(jan20041E, 101)
jan20041F <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "1F")
jan20041F <- sample_n(jan20041F, 188)
jan20042G <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "2G")
jan20042G <- sample_n(jan20042G, 335) 
jan20042H <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "2H")
jan20042H <- sample_n(jan20042H, 355) 
jan20042J <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "2J")
jan20042J <- sample_n(jan20042J, 419) 
jan20043k <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "3K")
jan20043k <- sample_n(jan20043k, 435) 
jan20043L <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "3L")
jan20043L <- sample_n(jan20043L, 366)
jan20043M <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "3M")
jan20043M <- sample_n(jan20043M, 23)
jan20043N <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "3N")
jan20043N <- sample_n(jan20043N, 132) 
jan20043o <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "3O")
jan20043o <- sample_n(jan20043o, 191) 
jan20043pn <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "3Pn")
jan20043pn <- sample_n(jan20043pn, 17) 
jan20043ps <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "3Ps")
jan20043ps <- sample_n(jan20043ps, 197) 
jan20044r <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "4R")
jan20044r <- sample_n(jan20044r, 95) 
jan20044s <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "4S")
jan20044s <- sample_n(jan20044s, 181) 
jan20044t <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "4T")
jan20044t <- sample_n(jan20044t, 251) 
jan20044vn <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "4Vn")
jan20044vn <- sample_n(jan20044vn, 52) 
jan20044vs <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "4Vs")
jan20044vs <- sample_n(jan20044vs, 197) 
jan20044w <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "4W")
jan20044w <- sample_n(jan20044w, 199) 
jan20044x <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "4X")
jan20044x <- sample_n(jan20044x, 187) 
jan20045y <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "5Y")
jan20045y <- sample_n(jan20045y, 74) 
jan20045ze <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "5Ze")
jan20045ze <- sample_n(jan20045ze, 24) 
jan2004hud <- subset(backno1, year == "2004" & month == "1" & nafo_zone == "HudsonStrait")
jan2004hud <- sample_n(jan2004hud, 197) 
jan2004 <- rbind(jan20040a, jan20040b, jan20041B, jan20041C, jan20041d, jan20041E, jan20041F,  jan20042G, jan20042H, jan20042J, jan20043k, jan20043L, jan20043M, jan20043N, jan20043o, jan20043pn, jan20043ps, jan20044r, jan20044s, jan20044t, jan20044vn, jan20044vs, jan20044w, jan20044x, jan20045y, jan20045ze, jan2004hud)
```


```{r jan 2005}
jan20050a <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "0A")
jan20050a <- sample_n(jan20050a, 16) 
jan20050b <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "0B")
jan20050b <- sample_n(jan20050b, 68) 
jan20051B <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "1B")
jan20051B <- sample_n(jan20051B, 1)
jan20051C <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "1C")
jan20051C <- sample_n(jan20051C, 6) 
jan20051d <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "1D")
jan20051d <- sample_n(jan20051d, 11)
jan20051E <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "1E")
jan20051E <- sample_n(jan20051E, 14)
jan20051F <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "1F")
jan20051F <- sample_n(jan20051F, 27)
jan20052G <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "2G")
jan20052G <- sample_n(jan20052G, 48) 
jan20052H <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "2H")
jan20052H <- sample_n(jan20052H, 50) 
jan20052J <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "2J")
jan20052J <- sample_n(jan20052J, 59) 
jan20053k <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "3K")
jan20053k <- sample_n(jan20053k, 62) 
jan20053L <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "3L")
jan20053L <- sample_n(jan20053L, 52)
jan20053M <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "3M")
jan20053M <- sample_n(jan20053M, 3)
jan20053N <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "3N")
jan20053N <- sample_n(jan20053N, 19) 
jan20053o <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "3O")
jan20053o <- sample_n(jan20053o, 27) 
jan20053pn <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "3Pn")
jan20053pn <- sample_n(jan20053pn, 2) 
jan20053ps <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "3Ps")
jan20053ps <- sample_n(jan20053ps, 28) 
jan20054r <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "4R")
jan20054r <- sample_n(jan20054r, 13) 
jan20054s <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "4S")
jan20054s <- sample_n(jan20054s, 26) 
jan20054t <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "4T")
jan20054t <- sample_n(jan20054t, 36) 
jan20054vn <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "4Vn")
jan20054vn <- sample_n(jan20054vn, 7) 
jan20054vs <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "4Vs")
jan20054vs <- sample_n(jan20054vs, 28) 
jan20054w <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "4W")
jan20054w <- sample_n(jan20054w, 28) 
jan20054x <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "4X")
jan20054x <- sample_n(jan20054x, 27) 
jan20055y <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "5Y")
jan20055y <- sample_n(jan20055y, 11) 
jan20055ze <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "5Ze")
jan20055ze <- sample_n(jan20055ze, 3) 
jan2005hud <- subset(backno1, year == "2005" & month == "1" & nafo_zone == "HudsonStrait")
jan2005hud <- sample_n(jan2005hud, 28) 
jan2005 <- rbind(jan20050a, jan20050b, jan20051B, jan20051C, jan20051d, jan20051E, jan20051F,  jan20052G, jan20052H, jan20052J, jan20053k, jan20053L, jan20053M, jan20053N, jan20053o, jan20053pn, jan20053ps, jan20054r, jan20054s, jan20054t, jan20054vn, jan20054vs, jan20054w, jan20054x, jan20055y, jan20055ze, jan2005hud)
```

```{r jan 2006}
jan20060a <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "0A")
jan20060a <- sample_n(jan20060a, 51) 
jan20060b <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "0B")
jan20060b <- sample_n(jan20060b, 207) 
jan20061B <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "1B")
jan20061B <- sample_n(jan20061B, 4)
jan20061C <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "1C")
jan20061C <- sample_n(jan20061C, 18) 
jan20061d <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "1D")
jan20061d <- sample_n(jan20061d, 33)
jan20061E <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "1E")
jan20061E <- sample_n(jan20061E, 44)
jan20061F <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "1F")
jan20061F <- sample_n(jan20061F, 82)
jan20062G <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "2G")
jan20062G <- sample_n(jan20062G, 145) 
jan20062H <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "2H")
jan20062H <- sample_n(jan20062H, 154) 
jan20062J <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "2J")
jan20062J <- sample_n(jan20062J, 182) 
jan20063k <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "3K")
jan20063k <- sample_n(jan20063k, 189) 
jan20063L <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "3L")
jan20063L <- sample_n(jan20063L, 158)
jan20063M <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "3M")
jan20063M <- sample_n(jan20063M, 10)
jan20063N <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "3N")
jan20063N <- sample_n(jan20063N, 57) 
jan20063o <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "3O")
jan20063o <- sample_n(jan20063o, 83) 
jan20063pn <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "3Pn")
jan20063pn <- sample_n(jan20063pn, 7) 
jan20063ps <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "3Ps")
jan20063ps <- sample_n(jan20063ps, 85) 
jan20064r <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "4R")
jan20064r <- sample_n(jan20064r, 41) 
jan20064s <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "4S")
jan20064s <- sample_n(jan20064s, 78) 
jan20064t <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "4T")
jan20064t <- sample_n(jan20064t, 109) 
jan20064vn <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "4Vn")
jan20064vn <- sample_n(jan20064vn, 23) 
jan20064vs <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "4Vs")
jan20064vs <- sample_n(jan20064vs, 85) 
jan20064w <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "4W")
jan20064w <- sample_n(jan20064w, 86) 
jan20064x <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "4X")
jan20064x <- sample_n(jan20064x, 81) 
jan20065y <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "5Y")
jan20065y <- sample_n(jan20065y, 32) 
jan20065ze <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "5Ze")
jan20065ze <- sample_n(jan20065ze, 11) 
jan2006hud <- subset(backno1, year == "2006" & month == "1" & nafo_zone == "HudsonStrait")
jan2006hud <- sample_n(jan2006hud, 85) 
jan2006 <- rbind(jan20060a, jan20060b, jan20061B, jan20061C, jan20061d, jan20061E, jan20061F,  jan20062G, jan20062H, jan20062J, jan20063k, jan20063L, jan20063M, jan20063N, jan20063o, jan20063pn, jan20063ps, jan20064r, jan20064s, jan20064t, jan20064vn, jan20064vs, jan20064w, jan20064x, jan20065y, jan20065ze, jan2006hud)
```

```{r merge all jan}
jan <- rbind(jan2003, jan2004, jan2005, jan2006)
```




```{r feb 2001}
feb20010a <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "0A")
feb20010a <- sample_n(feb20010a, 118) 
feb20010b <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "0B")
feb20010b <- sample_n(feb20010b, 485) 
feb20011B <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "1B")
feb20011B <- sample_n(feb20011B, 9)
feb20011C <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "1C")
feb20011C <- sample_n(feb20011C, 42) 
feb20011d <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "1D")
feb20011d <- sample_n(feb20011d, 78)
feb20011E <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "1E")
feb20011E <- sample_n(feb20011E, 102)
feb20011F <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "1F")
feb20011F <- sample_n(feb20011F, 191)
feb20012G <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "2G")
feb20012G <- sample_n(feb20012G, 340) 
feb20012H <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "2H")
feb20012H <- sample_n(feb20012H, 359) 
feb20012J <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "2J")
feb20012J <- sample_n(feb20012J, 424) 
feb20013k <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "3K")
feb20013k <- sample_n(feb20013k, 440) 
feb20013L <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "3L")
feb20013L <- sample_n(feb20013L, 370)
feb20013M <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "3M")
feb20013M <- sample_n(feb20013M, 23)
feb20013N <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "3N")
feb20013N <- sample_n(feb20013N, 134) 
feb20013o <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "3O")
feb20013o <- sample_n(feb20013o, 193) 
feb20013pn <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "3Pn")
feb20013pn <- sample_n(feb20013pn, 17) 
feb20013ps <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "3Ps")
feb20013ps <- sample_n(feb20013ps, 199) 
feb20014r <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "4R")
feb20014r <- sample_n(feb20014r, 96) 
feb20014s <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "4S")
feb20014s <- sample_n(feb20014s, 183) 
feb20014t <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "4T")
feb20014t <- sample_n(feb20014t, 255) 
feb20014vn <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "4Vn")
feb20014vn <- sample_n(feb20014vn, 53) 
feb20014vs <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "4Vs")
feb20014vs <- sample_n(feb20014vs, 199) 
feb20014w <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "4W")
feb20014w <- sample_n(feb20014w, 202) 
feb20014x <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "4X")
feb20014x <- sample_n(feb20014x, 189) 
feb20015y <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "5Y")
feb20015y <- sample_n(feb20015y, 75) 
feb20015ze <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "5Ze")
feb20015ze <- sample_n(feb20015ze, 25) 
feb2001hud <- subset(backno1, year == "2001" & month == "2" & nafo_zone == "HudsonStrait")
feb2001hud <- sample_n(feb2001hud, 199) 
feb2001 <- rbind(feb20010a, feb20010b, feb20011B, feb20011C, feb20011d, feb20011E, feb20011F,  feb20012G, feb20012H, feb20012J, feb20013k, feb20013L, feb20013M, feb20013N, feb20013o, feb20013pn, feb20013ps, feb20014r, feb20014s, feb20014t, feb20014vn, feb20014vs, feb20014w, feb20014x, feb20015y, feb20015ze, feb2001hud)
```


```{r feb 2005}
feb20050a <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "0A")
feb20050a <- sample_n(feb20050a, 94) 
feb20050b <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "0B")
feb20050b <- sample_n(feb20050b, 388) 
feb20051B <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "1B")
feb20051B <- sample_n(feb20051B, 7)
feb20051C <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "1C")
feb20051C <- sample_n(feb20051C, 33) 
feb20051d <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "1D")
feb20051d <- sample_n(feb20051d, 62)
feb20051E <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "1E")
feb20051E <- sample_n(feb20051E, 82)
feb20051F <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "1F")
feb20051F <- sample_n(feb20051F, 153)
feb20052G <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "2G")
feb20052G <- sample_n(feb20052G, 272) 
feb20052H <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "2H")
feb20052H <- sample_n(feb20052H, 287) 
feb20052J <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "2J")
feb20052J <- sample_n(feb20052J, 339) 
feb20053k <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "3K")
feb20053k <- sample_n(feb20053k, 352) 
feb20053L <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "3L")
feb20053L <- sample_n(feb20053L, 296)
feb20053M <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "3M")
feb20053M <- sample_n(feb20053M, 19)
feb20053N <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "3N")
feb20053N <- sample_n(feb20053N, 107) 
feb20053o <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "3O")
feb20053o <- sample_n(feb20053o, 155) 
feb20053pn <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "3Pn")
feb20053pn <- sample_n(feb20053pn, 14) 
feb20053ps <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "3Ps")
feb20053ps <- sample_n(feb20053ps, 159) 
feb20054r <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "4R")
feb20054r <- sample_n(feb20054r, 77) 
feb20054s <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "4S")
feb20054s <- sample_n(feb20054s, 147) 
feb20054t <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "4T")
feb20054t <- sample_n(feb20054t, 204) 
feb20054vn <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "4Vn")
feb20054vn <- sample_n(feb20054vn, 42) 
feb20054vs <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "4Vs")
feb20054vs <- sample_n(feb20054vs, 159) 
feb20054w <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "4W")
feb20054w <- sample_n(feb20054w, 161) 
feb20054x <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "4X")
feb20054x <- sample_n(feb20054x, 152) 
feb20055y <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "5Y")
feb20055y <- sample_n(feb20055y, 60) 
feb20055ze <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "5Ze")
feb20055ze <- sample_n(feb20055ze, 20) 
feb2005hud <- subset(backno1, year == "2005" & month == "2" & nafo_zone == "HudsonStrait")
feb2005hud <- sample_n(feb2005hud, 159) 
feb2005 <- rbind(feb20050a, feb20050b, feb20051B, feb20051C, feb20051d, feb20051E, feb20051F,  feb20052G, feb20052H, feb20052J, feb20053k, feb20053L, feb20053M, feb20053N, feb20053o, feb20053pn, feb20053ps, feb20054r, feb20054s, feb20054t, feb20054vn, feb20054vs, feb20054w, feb20054x, feb20055y, feb20055ze, feb2005hud)
```

```{r feb 2006}
feb20060a <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "0A")
feb20060a <- sample_n(feb20060a, 23) 
feb20060b <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "0B")
feb20060b <- sample_n(feb20060b, 97) 
feb20061B <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "1B")
feb20061B <- sample_n(feb20061B, 2)
feb20061C <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "1C")
feb20061C <- sample_n(feb20061C, 8) 
feb20061d <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "1D")
feb20061d <- sample_n(feb20061d, 15)
feb20061E <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "1E")
feb20061E <- sample_n(feb20061E, 20)
feb20061F <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "1F")
feb20061F <- sample_n(feb20061F, 38)
feb20062G <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "2G")
feb20062G <- sample_n(feb20062G, 68) 
feb20062H <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "2H")
feb20062H <- sample_n(feb20062H, 72) 
feb20062J <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "2J")
feb20062J <- sample_n(feb20062J, 85) 
feb20063k <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "3K")
feb20063k <- sample_n(feb20063k, 88) 
feb20063L <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "3L")
feb20063L <- sample_n(feb20063L, 74)
feb20063M <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "3M")
feb20063M <- sample_n(feb20063M, 5)
feb20063N <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "3N")
feb20063N <- sample_n(feb20063N, 27) 
feb20063o <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "3O")
feb20063o <- sample_n(feb20063o, 39) 
feb20063pn <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "3Pn")
feb20063pn <- sample_n(feb20063pn, 3) 
feb20063ps <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "3Ps")
feb20063ps <- sample_n(feb20063ps, 40) 
feb20064r <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "4R")
feb20064r <- sample_n(feb20064r, 19) 
feb20064s <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "4S")
feb20064s <- sample_n(feb20064s, 37) 
feb20064t <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "4T")
feb20064t <- sample_n(feb20064t, 51) 
feb20064vn <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "4Vn")
feb20064vn <- sample_n(feb20064vn, 11) 
feb20064vs <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "4Vs")
feb20064vs <- sample_n(feb20064vs, 40) 
feb20064w <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "4W")
feb20064w <- sample_n(feb20064w, 40) 
feb20064x <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "4X")
feb20064x <- sample_n(feb20064x, 38) 
feb20065y <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "5Y")
feb20065y <- sample_n(feb20065y, 15) 
feb20065ze <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "5Ze")
feb20065ze <- sample_n(feb20065ze, 5) 
feb2006hud <- subset(backno1, year == "2006" & month == "2" & nafo_zone == "HudsonStrait")
feb2006hud <- sample_n(feb2006hud, 40) 
feb2006 <- rbind(feb20060a, feb20060b, feb20061B, feb20061C, feb20061d, feb20061E, feb20061F,  feb20062G, feb20062H, feb20062J, feb20063k, feb20063L, feb20063M, feb20063N, feb20063o, feb20063pn, feb20063ps, feb20064r, feb20064s, feb20064t, feb20064vn, feb20064vs, feb20064w, feb20064x, feb20065y, feb20065ze, feb2006hud)
```

```{r merge all feb}
feb <- rbind(feb2001, feb2005, feb2006)
```



```{r mar 1999}
mar19990a <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "0A")
mar19990a <- sample_n(mar19990a, 34) 
mar19990b <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "0B")
mar19990b <- sample_n(mar19990b, 139) 
mar19991B <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "1B")
mar19991B <- sample_n(mar19991B, 2)
mar19991C <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "1C")
mar19991C <- sample_n(mar19991C, 12) 
mar19991d <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "1D")
mar19991d <- sample_n(mar19991d, 22)
mar19991E <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "1E")
mar19991E <- sample_n(mar19991E, 29)
mar19991F <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "1F")
mar19991F <- sample_n(mar19991F, 55)
mar19992G <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "2G")
mar19992G <- sample_n(mar19992G, 97) 
mar19992H <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "2H")
mar19992H <- sample_n(mar19992H, 103) 
mar19992J <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "2J")
mar19992J <- sample_n(mar19992J, 122) 
mar19993k <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "3K")
mar19993k <- sample_n(mar19993k, 126) 
mar19993L <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "3L")
mar19993L <- sample_n(mar19993L, 106)
mar19993M <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "3M")
mar19993M <- sample_n(mar19993M, 7)
mar19993N <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "3N")
mar19993N <- sample_n(mar19993N, 38) 
mar19993o <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "3O")
mar19993o <- sample_n(mar19993o, 55) 
mar19993pn <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "3Pn")
mar19993pn <- sample_n(mar19993pn, 5) 
mar19993ps <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "3Ps")
mar19993ps <- sample_n(mar19993ps, 57) 
mar19994r <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "4R")
mar19994r <- sample_n(mar19994r, 28) 
mar19994s <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "4S")
mar19994s <- sample_n(mar19994s, 53) 
mar19994t <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "4T")
mar19994t <- sample_n(mar19994t, 73) 
mar19994vn <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "4Vn")
mar19994vn <- sample_n(mar19994vn, 15) 
mar19994vs <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "4Vs")
mar19994vs <- sample_n(mar19994vs, 57) 
mar19994w <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "4W")
mar19994w <- sample_n(mar19994w, 58) 
mar19994x <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "4X")
mar19994x <- sample_n(mar19994x, 54) 
mar19995y <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "5Y")
mar19995y <- sample_n(mar19995y, 22) 
mar19995ze <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "5Ze")
mar19995ze <- sample_n(mar19995ze, 7) 
mar1999hud <- subset(backno1, year == "1999" & month == "3" & nafo_zone == "HudsonStrait")
mar1999hud <- sample_n(mar1999hud, 57) 
mar1999 <- rbind(mar19990a, mar19990b, mar19991B, mar19991C, mar19991d, mar19991E, mar19991F,  mar19992G, mar19992H, mar19992J, mar19993k, mar19993L, mar19993M, mar19993N, mar19993o, mar19993pn, mar19993ps, mar19994r, mar19994s, mar19994t, mar19994vn, mar19994vs, mar19994w, mar19994x, mar19995y, mar19995ze, mar1999hud)
```


```{r mar 2000}
mar20000a <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "0A")
mar20000a <- sample_n(mar20000a, 30) 
mar20000b <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "0B")
mar20000b <- sample_n(mar20000b, 124) 
mar20001B <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "1B")
mar20001B <- sample_n(mar20001B, 2)
mar20001C <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "1C")
mar20001C <- sample_n(mar20001C, 11) 
mar20001d <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "1D")
mar20001d <- sample_n(mar20001d, 20)
mar20001E <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "1E")
mar20001E <- sample_n(mar20001E, 26)
mar20001F <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "1F")
mar20001F <- sample_n(mar20001F, 49)
mar20002G <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "2G")
mar20002G <- sample_n(mar20002G, 87) 
mar20002H <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "2H")
mar20002H <- sample_n(mar20002H, 92) 
mar20002J <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "2J")
mar20002J <- sample_n(mar20002J, 108) 
mar20003k <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "3K")
mar20003k <- sample_n(mar20003k, 112) 
mar20003L <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "3L")
mar20003L <- sample_n(mar20003L, 95)
mar20003M <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "3M")
mar20003M <- sample_n(mar20003M, 6)
mar20003N <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "3N")
mar20003N <- sample_n(mar20003N, 34) 
mar20003o <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "3O")
mar20003o <- sample_n(mar20003o, 49) 
mar20003pn <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "3Pn")
mar20003pn <- sample_n(mar20003pn, 4) 
mar20003ps <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "3Ps")
mar20003ps <- sample_n(mar20003ps, 51) 
mar20004r <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "4R")
mar20004r <- sample_n(mar20004r, 25) 
mar20004s <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "4S")
mar20004s <- sample_n(mar20004s, 47) 
mar20004t <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "4T")
mar20004t <- sample_n(mar20004t, 65) 
mar20004vn <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "4Vn")
mar20004vn <- sample_n(mar20004vn, 14) 
mar20004vs <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "4Vs")
mar20004vs <- sample_n(mar20004vs, 51) 
mar20004w <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "4W")
mar20004w <- sample_n(mar20004w, 51) 
mar20004x <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "4X")
mar20004x <- sample_n(mar20004x, 48) 
mar20005y <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "5Y")
mar20005y <- sample_n(mar20005y, 19) 
mar20005ze <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "5Ze")
mar20005ze <- sample_n(mar20005ze, 6) 
mar2000hud <- subset(backno1, year == "2000" & month == "3" & nafo_zone == "HudsonStrait")
mar2000hud <- sample_n(mar2000hud, 51) 
mar2000 <- rbind(mar20000a, mar20000b, mar20001B, mar20001C, mar20001d, mar20001E, mar20001F,  mar20002G, mar20002H, mar20002J, mar20003k, mar20003L, mar20003M, mar20003N, mar20003o, mar20003pn, mar20003ps, mar20004r, mar20004s, mar20004t, mar20004vn, mar20004vs, mar20004w, mar20004x, mar20005y, mar20005ze, mar2000hud)
```


```{r mar 2001}
mar20010a <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "0A")
mar20010a <- sample_n(mar20010a, 18) 
mar20010b <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "0B")
mar20010b <- sample_n(mar20010b, 72) 
mar20011B <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "1B")
mar20011B <- sample_n(mar20011B, 1)
mar20011C <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "1C")
mar20011C <- sample_n(mar20011C, 6) 
mar20011d <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "1D")
mar20011d <- sample_n(mar20011d, 12)
mar20011E <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "1E")
mar20011E <- sample_n(mar20011E, 15)
mar20011F <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "1F")
mar20011F <- sample_n(mar20011F, 29)
mar20012G <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "2G")
mar20012G <- sample_n(mar20012G, 51) 
mar20012H <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "2H")
mar20012H <- sample_n(mar20012H, 54) 
mar20012J <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "2J")
mar20012J <- sample_n(mar20012J, 63) 
mar20013k <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "3K")
mar20013k <- sample_n(mar20013k, 66) 
mar20013L <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "3L")
mar20013L <- sample_n(mar20013L, 55)
mar20013M <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "3M")
mar20013M <- sample_n(mar20013M, 4)
mar20013N <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "3N")
mar20013N <- sample_n(mar20013N, 20) 
mar20013o <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "3O")
mar20013o <- sample_n(mar20013o, 29) 
mar20013pn <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "3Pn")
mar20013pn <- sample_n(mar20013pn, 3) 
mar20013ps <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "3Ps")
mar20013ps <- sample_n(mar20013ps, 30) 
mar20014r <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "4R")
mar20014r <- sample_n(mar20014r, 14) 
mar20014s <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "4S")
mar20014s <- sample_n(mar20014s, 27) 
mar20014t <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "4T")
mar20014t <- sample_n(mar20014t, 38) 
mar20014vn <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "4Vn")
mar20014vn <- sample_n(mar20014vn, 8) 
mar20014vs <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "4Vs")
mar20014vs <- sample_n(mar20014vs, 30) 
mar20014w <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "4W")
mar20014w <- sample_n(mar20014w, 30) 
mar20014x <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "4X")
mar20014x <- sample_n(mar20014x, 28) 
mar20015y <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "5Y")
mar20015y <- sample_n(mar20015y, 11) 
mar20015ze <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "5Ze")
mar20015ze <- sample_n(mar20015ze, 4) 
mar2001hud <- subset(backno1, year == "2001" & month == "3" & nafo_zone == "HudsonStrait")
mar2001hud <- sample_n(mar2001hud, 30) 
mar2001 <- rbind(mar20010a, mar20010b, mar20011B, mar20011C, mar20011d, mar20011E, mar20011F,  mar20012G, mar20012H, mar20012J, mar20013k, mar20013L, mar20013M, mar20013N, mar20013o, mar20013pn, mar20013ps, mar20014r, mar20014s, mar20014t, mar20014vn, mar20014vs, mar20014w, mar20014x, mar20015y, mar20015ze, mar2001hud)
```

```{r mar 2002}
mar20020a <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "0A")
mar20020a <- sample_n(mar20020a, 24) 
mar20020b <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "0B")
mar20020b <- sample_n(mar20020b, 96) 
mar20021B <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "1B")
mar20021B <- sample_n(mar20021B, 2)
mar20021C <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "1C")
mar20021C <- sample_n(mar20021C, 8) 
mar20021d <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "1D")
mar20021d <- sample_n(mar20021d, 15)
mar20021E <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "1E")
mar20021E <- sample_n(mar20021E, 20)
mar20021F <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "1F")
mar20021F <- sample_n(mar20021F, 38)
mar20022G <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "2G")
mar20022G <- sample_n(mar20022G, 68) 
mar20022H <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "2H")
mar20022H <- sample_n(mar20022H, 72) 
mar20022J <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "2J")
mar20022J <- sample_n(mar20022J, 84) 
mar20023k <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "3K")
mar20023k <- sample_n(mar20023k, 88) 
mar20023L <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "3L")
mar20023L <- sample_n(mar20023L, 74)
mar20023M <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "3M")
mar20023M <- sample_n(mar20023M, 5)
mar20023N <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "3N")
mar20023N <- sample_n(mar20023N, 27) 
mar20023o <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "3O")
mar20023o <- sample_n(mar20023o, 38) 
mar20023pn <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "3Pn")
mar20023pn <- sample_n(mar20023pn, 3) 
mar20023ps <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "3Ps")
mar20023ps <- sample_n(mar20023ps, 40) 
mar20024r <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "4R")
mar20024r <- sample_n(mar20024r, 19) 
mar20024s <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "4S")
mar20024s <- sample_n(mar20024s, 36) 
mar20024t <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "4T")
mar20024t <- sample_n(mar20024t, 51) 
mar20024vn <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "4Vn")
mar20024vn <- sample_n(mar20024vn, 10) 
mar20024vs <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "4Vs")
mar20024vs <- sample_n(mar20024vs, 40) 
mar20024w <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "4W")
mar20024w <- sample_n(mar20024w, 40) 
mar20024x <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "4X")
mar20024x <- sample_n(mar20024x, 38) 
mar20025y <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "5Y")
mar20025y <- sample_n(mar20025y, 15) 
mar20025ze <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "5Ze")
mar20025ze <- sample_n(mar20025ze, 5) 
mar2002hud <- subset(backno1, year == "2002" & month == "3" & nafo_zone == "HudsonStrait")
mar2002hud <- sample_n(mar2002hud, 40) 
mar2002 <- rbind(mar20020a, mar20020b, mar20021B, mar20021C, mar20021d, mar20021E, mar20021F,  mar20022G, mar20022H, mar20022J, mar20023k, mar20023L, mar20023M, mar20023N, mar20023o, mar20023pn, mar20023ps, mar20024r, mar20024s, mar20024t, mar20024vn, mar20024vs, mar20024w, mar20024x, mar20025y, mar20025ze, mar2002hud)
```

```{r mar 2003}
mar20030a <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "0A")
mar20030a <- sample_n(mar20030a, 17) 
mar20030b <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "0B")
mar20030b <- sample_n(mar20030b, 68) 
mar20031B <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "1B")
mar20031B <- sample_n(mar20031B, 1)
mar20031C <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "1C")
mar20031C <- sample_n(mar20031C, 6) 
mar20031d <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "1D")
mar20031d <- sample_n(mar20031d, 11)
mar20031E <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "1E")
mar20031E <- sample_n(mar20031E, 15)
mar20031F <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "1F")
mar20031F <- sample_n(mar20031F, 27)
mar20032G <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "2G")
mar20032G <- sample_n(mar20032G, 49) 
mar20032H <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "2H")
mar20032H <- sample_n(mar20032H, 51) 
mar20032J <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "2J")
mar20032J <- sample_n(mar20032J, 61) 
mar20033k <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "3K")
mar20033k <- sample_n(mar20033k, 63) 
mar20033L <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "3L")
mar20033L <- sample_n(mar20033L, 53)
mar20033M <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "3M")
mar20033M <- sample_n(mar20033M, 3)
mar20033N <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "3N")
mar20033N <- sample_n(mar20033N, 19) 
mar20033o <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "3O")
mar20033o <- sample_n(mar20033o, 28) 
mar20033pn <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "3Pn")
mar20033pn <- sample_n(mar20033pn, 2) 
mar20033ps <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "3Ps")
mar20033ps <- sample_n(mar20033ps, 29) 
mar20034r <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "4R")
mar20034r <- sample_n(mar20034r, 14) 
mar20034s <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "4S")
mar20034s <- sample_n(mar20034s, 26) 
mar20034t <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "4T")
mar20034t <- sample_n(mar20034t, 37) 
mar20034vn <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "4Vn")
mar20034vn <- sample_n(mar20034vn, 8) 
mar20034vs <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "4Vs")
mar20034vs <- sample_n(mar20034vs, 29) 
mar20034w <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "4W")
mar20034w <- sample_n(mar20034w, 29) 
mar20034x <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "4X")
mar20034x <- sample_n(mar20034x, 27) 
mar20035y <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "5Y")
mar20035y <- sample_n(mar20035y, 11) 
mar20035ze <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "5Ze")
mar20035ze <- sample_n(mar20035ze, 4) 
mar2003hud <- subset(backno1, year == "2003" & month == "3" & nafo_zone == "HudsonStrait")
mar2003hud <- sample_n(mar2003hud, 29) 
mar2003 <- rbind(mar20030a, mar20030b, mar20031B, mar20031C, mar20031d, mar20031E, mar20031F,  mar20032G, mar20032H, mar20032J, mar20033k, mar20033L, mar20033M, mar20033N, mar20033o, mar20033pn, mar20033ps, mar20034r, mar20034s, mar20034t, mar20034vn, mar20034vs, mar20034w, mar20034x, mar20035y, mar20035ze, mar2003hud)
```

```{r mar 2005}
mar20050a <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "0A")
mar20050a <- sample_n(mar20050a, 21) 
mar20050b <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "0B")
mar20050b <- sample_n(mar20050b, 84) 
mar20051B <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "1B")
mar20051B <- sample_n(mar20051B, 2)
mar20051C <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "1C")
mar20051C <- sample_n(mar20051C, 7) 
mar20051d <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "1D")
mar20051d <- sample_n(mar20051d, 13)
mar20051E <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "1E")
mar20051E <- sample_n(mar20051E, 18)
mar20051F <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "1F")
mar20051F <- sample_n(mar20051F, 33)
mar20052G <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "2G")
mar20052G <- sample_n(mar20052G, 59) 
mar20052H <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "2H")
mar20052H <- sample_n(mar20052H, 63) 
mar20052J <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "2J")
mar20052J <- sample_n(mar20052J, 74) 
mar20053k <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "3K")
mar20053k <- sample_n(mar20053k, 77) 
mar20053L <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "3L")
mar20053L <- sample_n(mar20053L, 65)
mar20053M <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "3M")
mar20053M <- sample_n(mar20053M, 4)
mar20053N <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "3N")
mar20053N <- sample_n(mar20053N, 23) 
mar20053o <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "3O")
mar20053o <- sample_n(mar20053o, 34) 
mar20053pn <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "3Pn")
mar20053pn <- sample_n(mar20053pn, 3) 
mar20053ps <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "3Ps")
mar20053ps <- sample_n(mar20053ps, 35) 
mar20054r <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "4R")
mar20054r <- sample_n(mar20054r, 17) 
mar20054s <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "4S")
mar20054s <- sample_n(mar20054s, 32) 
mar20054t <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "4T")
mar20054t <- sample_n(mar20054t, 44) 
mar20054vn <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "4Vn")
mar20054vn <- sample_n(mar20054vn, 9) 
mar20054vs <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "4Vs")
mar20054vs <- sample_n(mar20054vs, 35) 
mar20054w <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "4W")
mar20054w <- sample_n(mar20054w, 35) 
mar20054x <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "4X")
mar20054x <- sample_n(mar20054x, 33) 
mar20055y <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "5Y")
mar20055y <- sample_n(mar20055y, 13) 
mar20055ze <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "5Ze")
mar20055ze <- sample_n(mar20055ze, 4) 
mar2005hud <- subset(backno1, year == "2005" & month == "3" & nafo_zone == "HudsonStrait")
mar2005hud <- sample_n(mar2005hud, 35) 
mar2005 <- rbind(mar20050a, mar20050b, mar20051B, mar20051C, mar20051d, mar20051E, mar20051F,  mar20052G, mar20052H, mar20052J, mar20053k, mar20053L, mar20053M, mar20053N, mar20053o, mar20053pn, mar20053ps, mar20054r, mar20054s, mar20054t, mar20054vn, mar20054vs, mar20054w, mar20054x, mar20055y, mar20055ze, mar2005hud)
```

```{r mar 2006}
mar20060a <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "0A")
mar20060a <- sample_n(mar20060a, 37) 
mar20060b <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "0B")
mar20060b <- sample_n(mar20060b, 151) 
mar20061B <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "1B")
mar20061B <- sample_n(mar20061B, 3)
mar20061C <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "1C")
mar20061C <- sample_n(mar20061C, 13) 
mar20061d <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "1D")
mar20061d <- sample_n(mar20061d, 24)
mar20061E <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "1E")
mar20061E <- sample_n(mar20061E, 32)
mar20061F <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "1F")
mar20061F <- sample_n(mar20061F, 59)
mar20062G <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "2G")
mar20062G <- sample_n(mar20062G, 106) 
mar20062H <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "2H")
mar20062H <- sample_n(mar20062H, 112) 
mar20062J <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "2J")
mar20062J <- sample_n(mar20062J, 132) 
mar20063k <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "3K")
mar20063k <- sample_n(mar20063k, 137) 
mar20063L <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "3L")
mar20063L <- sample_n(mar20063L, 115)
mar20063M <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "3M")
mar20063M <- sample_n(mar20063M, 7)
mar20063N <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "3N")
mar20063N <- sample_n(mar20063N, 42) 
mar20063o <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "3O")
mar20063o <- sample_n(mar20063o, 60) 
mar20063pn <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "3Pn")
mar20063pn <- sample_n(mar20063pn, 6) 
mar20063ps <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "3Ps")
mar20063ps <- sample_n(mar20063ps, 62) 
mar20064r <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "4R")
mar20064r <- sample_n(mar20064r, 30) 
mar20064s <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "4S")
mar20064s <- sample_n(mar20064s, 57) 
mar20064t <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "4T")
mar20064t <- sample_n(mar20064t, 79) 
mar20064vn <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "4Vn")
mar20064vn <- sample_n(mar20064vn, 17) 
mar20064vs <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "4Vs")
mar20064vs <- sample_n(mar20064vs, 62) 
mar20064w <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "4W")
mar20064w <- sample_n(mar20064w, 63) 
mar20064x <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "4X")
mar20064x <- sample_n(mar20064x, 59) 
mar20065y <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "5Y")
mar20065y <- sample_n(mar20065y, 23) 
mar20065ze <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "5Ze")
mar20065ze <- sample_n(mar20065ze, 8) 
mar2006hud <- subset(backno1, year == "2006" & month == "3" & nafo_zone == "HudsonStrait")
mar2006hud <- sample_n(mar2006hud, 62) 
mar2006 <- rbind(mar20060a, mar20060b, mar20061B, mar20061C, mar20061d, mar20061E, mar20061F,  mar20062G, mar20062H, mar20062J, mar20063k, mar20063L, mar20063M, mar20063N, mar20063o, mar20063pn, mar20063ps, mar20064r, mar20064s, mar20064t, mar20064vn, mar20064vs, mar20064w, mar20064x, mar20065y, mar20065ze, mar2006hud)
```


```{r mar 2007}
mar20070a <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "0A")
mar20070a <- sample_n(mar20070a, 14) 
mar20070b <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "0B")
mar20070b <- sample_n(mar20070b, 59) 
mar20071B <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "1B")
mar20071B <- sample_n(mar20071B, 1)
mar20071C <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "1C")
mar20071C <- sample_n(mar20071C, 5) 
mar20071d <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "1D")
mar20071d <- sample_n(mar20071d, 10)
mar20071E <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "1E")
mar20071E <- sample_n(mar20071E, 13)
mar20071F <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "1F")
mar20071F <- sample_n(mar20071F, 24)
mar20072G <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "2G")
mar20072G <- sample_n(mar20072G, 42) 
mar20072H <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "2H")
mar20072H <- sample_n(mar20072H, 45) 
mar20072J <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "2J")
mar20072J <- sample_n(mar20072J, 53) 
mar20073k <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "3K")
mar20073k <- sample_n(mar20073k, 55) 
mar20073L <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "3L")
mar20073L <- sample_n(mar20073L, 46)
mar20073M <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "3M")
mar20073M <- sample_n(mar20073M, 3)
mar20073N <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "3N")
mar20073N <- sample_n(mar20073N, 17) 
mar20073o <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "3O")
mar20073o <- sample_n(mar20073o, 24) 
mar20073pn <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "3Pn")
mar20073pn <- sample_n(mar20073pn, 2) 
mar20073ps <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "3Ps")
mar20073ps <- sample_n(mar20073ps, 25) 
mar20074r <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "4R")
mar20074r <- sample_n(mar20074r, 12) 
mar20074s <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "4S")
mar20074s <- sample_n(mar20074s, 23) 
mar20074t <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "4T")
mar20074t <- sample_n(mar20074t, 32) 
mar20074vn <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "4Vn")
mar20074vn <- sample_n(mar20074vn, 7) 
mar20074vs <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "4Vs")
mar20074vs <- sample_n(mar20074vs, 25) 
mar20074w <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "4W")
mar20074w <- sample_n(mar20074w, 25) 
mar20074x <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "4X")
mar20074x <- sample_n(mar20074x, 24) 
mar20075y <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "5Y")
mar20075y <- sample_n(mar20075y, 9) 
mar20075ze <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "5Ze")
mar20075ze <- sample_n(mar20075ze, 3) 
mar2007hud <- subset(backno1, year == "2007" & month == "3" & nafo_zone == "HudsonStrait")
mar2007hud <- sample_n(mar2007hud, 25) 
mar2007 <- rbind(mar20070a, mar20070b, mar20071B, mar20071C, mar20071d, mar20071E, mar20071F,  mar20072G, mar20072H, mar20072J, mar20073k, mar20073L, mar20073M, mar20073N, mar20073o, mar20073pn, mar20073ps, mar20074r, mar20074s, mar20074t, mar20074vn, mar20074vs, mar20074w, mar20074x, mar20075y, mar20075ze, mar2007hud)
```


```{r mar 2008}
mar20080a <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "0A")
mar20080a <- sample_n(mar20080a, 22) 
mar20080b <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "0B")
mar20080b <- sample_n(mar20080b, 91) 
mar20081B <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "1B")
mar20081B <- sample_n(mar20081B, 2)
mar20081C <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "1C")
mar20081C <- sample_n(mar20081C, 8) 
mar20081d <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "1D")
mar20081d <- sample_n(mar20081d, 15)
mar20081E <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "1E")
mar20081E <- sample_n(mar20081E, 19)
mar20081F <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "1F")
mar20081F <- sample_n(mar20081F, 36)
mar20082G <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "2G")
mar20082G <- sample_n(mar20082G, 63) 
mar20082H <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "2H")
mar20082H <- sample_n(mar20082H, 67) 
mar20082J <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "2J")
mar20082J <- sample_n(mar20082J, 79) 
mar20083k <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "3K")
mar20083k <- sample_n(mar20083k, 82) 
mar20083L <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "3L")
mar20083L <- sample_n(mar20083L, 69)
mar20083M <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "3M")
mar20083M <- sample_n(mar20083M, 4)
mar20083N <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "3N")
mar20083N <- sample_n(mar20083N, 25) 
mar20083o <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "3O")
mar20083o <- sample_n(mar20083o, 36) 
mar20083pn <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "3Pn")
mar20083pn <- sample_n(mar20083pn, 3) 
mar20083ps <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "3Ps")
mar20083ps <- sample_n(mar20083ps, 37) 
mar20084r <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "4R")
mar20084r <- sample_n(mar20084r, 18) 
mar20084s <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "4S")
mar20084s <- sample_n(mar20084s, 34) 
mar20084t <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "4T")
mar20084t <- sample_n(mar20084t, 48) 
mar20084vn <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "4Vn")
mar20084vn <- sample_n(mar20084vn, 10) 
mar20084vs <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "4Vs")
mar20084vs <- sample_n(mar20084vs, 37) 
mar20084w <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "4W")
mar20084w <- sample_n(mar20084w, 38) 
mar20084x <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "4X")
mar20084x <- sample_n(mar20084x, 35) 
mar20085y <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "5Y")
mar20085y <- sample_n(mar20085y, 14) 
mar20085ze <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "5Ze")
mar20085ze <- sample_n(mar20085ze, 5) 
mar2008hud <- subset(backno1, year == "2008" & month == "3" & nafo_zone == "HudsonStrait")
mar2008hud <- sample_n(mar2008hud, 37) 
mar2008 <- rbind(mar20080a, mar20080b, mar20081B, mar20081C, mar20081d, mar20081E, mar20081F,  mar20082G, mar20082H, mar20082J, mar20083k, mar20083L, mar20083M, mar20083N, mar20083o, mar20083pn, mar20083ps, mar20084r, mar20084s, mar20084t, mar20084vn, mar20084vs, mar20084w, mar20084x, mar20085y, mar20085ze, mar2008hud)
```

```{r mar 2009}
mar20090a <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "0A")
mar20090a <- sample_n(mar20090a, 7) 
mar20090b <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "0B")
mar20090b <- sample_n(mar20090b, 30) 
mar20091B <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "1B")
mar20091B <- sample_n(mar20091B, 1)
mar20091C <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "1C")
mar20091C <- sample_n(mar20091C, 3) 
mar20091d <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "1D")
mar20091d <- sample_n(mar20091d, 5)
mar20091E <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "1E")
mar20091E <- sample_n(mar20091E, 6)
mar20091F <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "1F")
mar20091F <- sample_n(mar20091F, 12)
mar20092G <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "2G")
mar20092G <- sample_n(mar20092G, 21) 
mar20092H <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "2H")
mar20092H <- sample_n(mar20092H, 22) 
mar20092J <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "2J")
mar20092J <- sample_n(mar20092J, 27) 
mar20093k <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "3K")
mar20093k <- sample_n(mar20093k, 28) 
mar20093L <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "3L")
mar20093L <- sample_n(mar20093L, 23)
mar20093M <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "3M")
mar20093M <- sample_n(mar20093M, 2)
mar20093N <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "3N")
mar20093N <- sample_n(mar20093N, 8) 
mar20093o <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "3O")
mar20093o <- sample_n(mar20093o, 12) 
mar20093pn <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "3Pn")
mar20093pn <- sample_n(mar20093pn, 1) 
mar20093ps <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "3Ps")
mar20093ps <- sample_n(mar20093ps, 12) 
mar20094r <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "4R")
mar20094r <- sample_n(mar20094r, 6) 
mar20094s <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "4S")
mar20094s <- sample_n(mar20094s, 11) 
mar20094t <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "4T")
mar20094t <- sample_n(mar20094t, 16) 
mar20094vn <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "4Vn")
mar20094vn <- sample_n(mar20094vn, 3) 
mar20094vs <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "4Vs")
mar20094vs <- sample_n(mar20094vs, 12) 
mar20094w <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "4W")
mar20094w <- sample_n(mar20094w, 13) 
mar20094x <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "4X")
mar20094x <- sample_n(mar20094x, 12) 
mar20095y <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "5Y")
mar20095y <- sample_n(mar20095y, 5) 
mar20095ze <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "5Ze")
mar20095ze <- sample_n(mar20095ze, 2) 
mar2009hud <- subset(backno1, year == "2009" & month == "3" & nafo_zone == "HudsonStrait")
mar2009hud <- sample_n(mar2009hud, 12) 
mar2009 <- rbind(mar20090a, mar20090b, mar20091B, mar20091C, mar20091d, mar20091E, mar20091F,  mar20092G, mar20092H, mar20092J, mar20093k, mar20093L, mar20093M, mar20093N, mar20093o, mar20093pn, mar20093ps, mar20094r, mar20094s, mar20094t, mar20094vn, mar20094vs, mar20094w, mar20094x, mar20095y, mar20095ze, mar2009hud)
```

```{r mar 2010}
mar20100a <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "0A")
mar20100a <- sample_n(mar20100a, 13) 
mar20100b <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "0B")
mar20100b <- sample_n(mar20100b, 51) 
mar20101B <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "1B")
mar20101B <- sample_n(mar20101B, 1)
mar20101C <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "1C")
mar20101C <- sample_n(mar20101C, 4) 
mar20101d <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "1D")
mar20101d <- sample_n(mar20101d, 8)
mar20101E <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "1E")
mar20101E <- sample_n(mar20101E, 11)
mar20101F <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "1F")
mar20101F <- sample_n(mar20101F, 20)
mar20102G <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "2G")
mar20102G <- sample_n(mar20102G, 36) 
mar20102H <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "2H")
mar20102H <- sample_n(mar20102H, 38) 
mar20102J <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "2J")
mar20102J <- sample_n(mar20102J, 45) 
mar20103k <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "3K")
mar20103k <- sample_n(mar20103k, 47) 
mar20103L <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "3L")
mar20103L <- sample_n(mar20103L, 39)
mar20103M <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "3M")
mar20103M <- sample_n(mar20103M, 3)
mar20103N <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "3N")
mar20103N <- sample_n(mar20103N, 14) 
mar20103o <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "3O")
mar20103o <- sample_n(mar20103o, 21) 
mar20103pn <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "3Pn")
mar20103pn <- sample_n(mar20103pn, 2) 
mar20103ps <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "3Ps")
mar20103ps <- sample_n(mar20103ps, 21) 
mar20104r <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "4R")
mar20104r <- sample_n(mar20104r, 10) 
mar20104s <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "4S")
mar20104s <- sample_n(mar20104s, 19) 
mar20104t <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "4T")
mar20104t <- sample_n(mar20104t, 27) 
mar20104vn <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "4Vn")
mar20104vn <- sample_n(mar20104vn, 6) 
mar20104vs <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "4Vs")
mar20104vs <- sample_n(mar20104vs, 21) 
mar20104w <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "4W")
mar20104w <- sample_n(mar20104w, 21) 
mar20104x <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "4X")
mar20104x <- sample_n(mar20104x, 20) 
mar20105y <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "5Y")
mar20105y <- sample_n(mar20105y, 8) 
mar20105ze <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "5Ze")
mar20105ze <- sample_n(mar20105ze, 3) 
mar2010hud <- subset(backno1, year == "2010" & month == "3" & nafo_zone == "HudsonStrait")
mar2010hud <- sample_n(mar2010hud, 21) 
mar2010 <- rbind(mar20100a, mar20100b, mar20101B, mar20101C, mar20101d, mar20101E, mar20101F,  mar20102G, mar20102H, mar20102J, mar20103k, mar20103L, mar20103M, mar20103N, mar20103o, mar20103pn, mar20103ps, mar20104r, mar20104s, mar20104t, mar20104vn, mar20104vs, mar20104w, mar20104x, mar20105y, mar20105ze, mar2010hud)
```

```{r merge all mar}
mar <- rbind(mar1999, mar2000, mar2001, mar2002, mar2003, mar2005, mar2006, mar2007, mar2008, mar2009, mar2010)
```


```{r apr 1998}
apr19980a <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "0A")
apr19980a <- sample_n(apr19980a, 20) 
apr19980b <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "0B")
apr19980b <- sample_n(apr19980b, 82) 
apr19981B <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "1B")
apr19981B <- sample_n(apr19981B, 1)
apr19981C <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "1C")
apr19981C <- sample_n(apr19981C, 7) 
apr19981d <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "1D")
apr19981d <- sample_n(apr19981d, 13)
apr19981E <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "1E")
apr19981E <- sample_n(apr19981E, 17)
apr19981F <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "1F")
apr19981F <- sample_n(apr19981F, 32)
apr19982G <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "2G")
apr19982G <- sample_n(apr19982G, 57) 
apr19982H <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "2H")
apr19982H <- sample_n(apr19982H, 61) 
apr19982J <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "2J")
apr19982J <- sample_n(apr19982J, 72) 
apr19983k <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "3K")
apr19983k <- sample_n(apr19983k, 74) 
apr19983L <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "3L")
apr19983L <- sample_n(apr19983L, 63)
apr19983M <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "3M")
apr19983M <- sample_n(apr19983M, 4)
apr19983N <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "3N")
apr19983N <- sample_n(apr19983N, 23) 
apr19983o <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "3O")
apr19983o <- sample_n(apr19983o, 33) 
apr19983pn <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "3Pn")
apr19983pn <- sample_n(apr19983pn, 3) 
apr19983ps <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "3Ps")
apr19983ps <- sample_n(apr19983ps, 34) 
apr19984r <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "4R")
apr19984r <- sample_n(apr19984r, 16) 
apr19984s <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "4S")
apr19984s <- sample_n(apr19984s, 31) 
apr19984t <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "4T")
apr19984t <- sample_n(apr19984t, 43) 
apr19984vn <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "4Vn")
apr19984vn <- sample_n(apr19984vn, 9) 
apr19984vs <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "4Vs")
apr19984vs <- sample_n(apr19984vs, 34) 
apr19984w <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "4W")
apr19984w <- sample_n(apr19984w, 34) 
apr19984x <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "4X")
apr19984x <- sample_n(apr19984x, 32) 
apr19985y <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "5Y")
apr19985y <- sample_n(apr19985y, 13) 
apr19985ze <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "5Ze")
apr19985ze <- sample_n(apr19985ze, 4) 
apr1998hud <- subset(backno1, year == "1998" & month == "4" & nafo_zone == "HudsonStrait")
apr1998hud <- sample_n(apr1998hud, 34) 
apr1998 <- rbind(apr19980a, apr19980b, apr19981B, apr19981C, apr19981d, apr19981E, apr19981F,  apr19982G, apr19982H, apr19982J, apr19983k, apr19983L, apr19983M, apr19983N, apr19983o, apr19983pn, apr19983ps, apr19984r, apr19984s, apr19984t, apr19984vn, apr19984vs, apr19984w, apr19984x, apr19985y, apr19985ze, apr1998hud)
```

```{r apr 1999}
apr19990a <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "0A")
apr19990a <- sample_n(apr19990a, 14) 
apr19990b <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "0B")
apr19990b <- sample_n(apr19990b, 59) 
apr19991B <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "1B")
apr19991B <- sample_n(apr19991B, 1)
apr19991C <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "1C")
apr19991C <- sample_n(apr19991C, 5) 
apr19991d <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "1D")
apr19991d <- sample_n(apr19991d, 9)
apr19991E <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "1E")
apr19991E <- sample_n(apr19991E, 12)
apr19991F <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "1F")
apr19991F <- sample_n(apr19991F, 23)
apr19992G <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "2G")
apr19992G <- sample_n(apr19992G, 41) 
apr19992H <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "2H")
apr19992H <- sample_n(apr19992H, 43) 
apr19992J <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "2J")
apr19992J <- sample_n(apr19992J, 51) 
apr19993k <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "3K")
apr19993k <- sample_n(apr19993k, 53) 
apr19993L <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "3L")
apr19993L <- sample_n(apr19993L, 45)
apr19993M <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "3M")
apr19993M <- sample_n(apr19993M, 3)
apr19993N <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "3N")
apr19993N <- sample_n(apr19993N, 16) 
apr19993o <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "3O")
apr19993o <- sample_n(apr19993o, 23) 
apr19993pn <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "3Pn")
apr19993pn <- sample_n(apr19993pn, 2) 
apr19993ps <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "3Ps")
apr19993ps <- sample_n(apr19993ps, 24) 
apr19994r <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "4R")
apr19994r <- sample_n(apr19994r, 12) 
apr19994s <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "4S")
apr19994s <- sample_n(apr19994s, 22) 
apr19994t <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "4T")
apr19994t <- sample_n(apr19994t, 31) 
apr19994vn <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "4Vn")
apr19994vn <- sample_n(apr19994vn, 7) 
apr19994vs <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "4Vs")
apr19994vs <- sample_n(apr19994vs, 24) 
apr19994w <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "4W")
apr19994w <- sample_n(apr19994w, 24) 
apr19994x <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "4X")
apr19994x <- sample_n(apr19994x, 23) 
apr19995y <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "5Y")
apr19995y <- sample_n(apr19995y, 9) 
apr19995ze <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "5Ze")
apr19995ze <- sample_n(apr19995ze, 3) 
apr1999hud <- subset(backno1, year == "1999" & month == "4" & nafo_zone == "HudsonStrait")
apr1999hud <- sample_n(apr1999hud, 24) 
apr1999 <- rbind(apr19990a, apr19990b, apr19991B, apr19991C, apr19991d, apr19991E, apr19991F,  apr19992G, apr19992H, apr19992J, apr19993k, apr19993L, apr19993M, apr19993N, apr19993o, apr19993pn, apr19993ps, apr19994r, apr19994s, apr19994t, apr19994vn, apr19994vs, apr19994w, apr19994x, apr19995y, apr19995ze, apr1999hud)
```


```{r apr 2000}
apr20000a <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "0A")
apr20000a <- sample_n(apr20000a, 14) 
apr20000b <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "0B")
apr20000b <- sample_n(apr20000b, 59) 
apr20001B <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "1B")
apr20001B <- sample_n(apr20001B, 1)
apr20001C <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "1C")
apr20001C <- sample_n(apr20001C, 5) 
apr20001d <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "1D")
apr20001d <- sample_n(apr20001d, 9)
apr20001E <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "1E")
apr20001E <- sample_n(apr20001E, 12)
apr20001F <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "1F")
apr20001F <- sample_n(apr20001F, 23)
apr20002G <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "2G")
apr20002G <- sample_n(apr20002G, 41) 
apr20002H <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "2H")
apr20002H <- sample_n(apr20002H, 43) 
apr20002J <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "2J")
apr20002J <- sample_n(apr20002J, 51) 
apr20003k <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "3K")
apr20003k <- sample_n(apr20003k, 53) 
apr20003L <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "3L")
apr20003L <- sample_n(apr20003L, 45)
apr20003M <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "3M")
apr20003M <- sample_n(apr20003M, 3)
apr20003N <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "3N")
apr20003N <- sample_n(apr20003N, 16) 
apr20003o <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "3O")
apr20003o <- sample_n(apr20003o, 23) 
apr20003pn <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "3Pn")
apr20003pn <- sample_n(apr20003pn, 2) 
apr20003ps <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "3Ps")
apr20003ps <- sample_n(apr20003ps, 24) 
apr20004r <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "4R")
apr20004r <- sample_n(apr20004r, 12) 
apr20004s <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "4S")
apr20004s <- sample_n(apr20004s, 22) 
apr20004t <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "4T")
apr20004t <- sample_n(apr20004t, 31) 
apr20004vn <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "4Vn")
apr20004vn <- sample_n(apr20004vn, 7) 
apr20004vs <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "4Vs")
apr20004vs <- sample_n(apr20004vs, 24) 
apr20004w <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "4W")
apr20004w <- sample_n(apr20004w, 24) 
apr20004x <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "4X")
apr20004x <- sample_n(apr20004x, 23) 
apr20005y <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "5Y")
apr20005y <- sample_n(apr20005y, 9) 
apr20005ze <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "5Ze")
apr20005ze <- sample_n(apr20005ze, 3) 
apr2000hud <- subset(backno1, year == "2000" & month == "4" & nafo_zone == "HudsonStrait")
apr2000hud <- sample_n(apr2000hud, 24) 
apr2000 <- rbind(apr20000a, apr20000b, apr20001B, apr20001C, apr20001d, apr20001E, apr20001F,  apr20002G, apr20002H, apr20002J, apr20003k, apr20003L, apr20003M, apr20003N, apr20003o, apr20003pn, apr20003ps, apr20004r, apr20004s, apr20004t, apr20004vn, apr20004vs, apr20004w, apr20004x, apr20005y, apr20005ze, apr2000hud)
```


```{r apr 2001}
apr20010a <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "0A")
apr20010a <- sample_n(apr20010a, 23) 
apr20010b <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "0B")
apr20010b <- sample_n(apr20010b, 96) 
apr20011B <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "1B")
apr20011B <- sample_n(apr20011B, 2)
apr20011C <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "1C")
apr20011C <- sample_n(apr20011C, 8) 
apr20011d <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "1D")
apr20011d <- sample_n(apr20011d, 15)
apr20011E <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "1E")
apr20011E <- sample_n(apr20011E, 20)
apr20011F <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "1F")
apr20011F <- sample_n(apr20011F, 38)
apr20012G <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "2G")
apr20012G <- sample_n(apr20012G, 67) 
apr20012H <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "2H")
apr20012H <- sample_n(apr20012H, 71) 
apr20012J <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "2J")
apr20012J <- sample_n(apr20012J, 84) 
apr20013k <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "3K")
apr20013k <- sample_n(apr20013k, 87) 
apr20013L <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "3L")
apr20013L <- sample_n(apr20013L, 73)
apr20013M <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "3M")
apr20013M <- sample_n(apr20013M, 5)
apr20013N <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "3N")
apr20013N <- sample_n(apr20013N, 27) 
apr20013o <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "3O")
apr20013o <- sample_n(apr20013o, 38) 
apr20013pn <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "3Pn")
apr20013pn <- sample_n(apr20013pn, 3) 
apr20013ps <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "3Ps")
apr20013ps <- sample_n(apr20013ps, 40) 
apr20014r <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "4R")
apr20014r <- sample_n(apr20014r, 19) 
apr20014s <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "4S")
apr20014s <- sample_n(apr20014s, 36) 
apr20014t <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "4T")
apr20014t <- sample_n(apr20014t, 50) 
apr20014vn <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "4Vn")
apr20014vn <- sample_n(apr20014vn, 11) 
apr20014vs <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "4Vs")
apr20014vs <- sample_n(apr20014vs, 40) 
apr20014w <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "4W")
apr20014w <- sample_n(apr20014w, 40) 
apr20014x <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "4X")
apr20014x <- sample_n(apr20014x, 38) 
apr20015y <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "5Y")
apr20015y <- sample_n(apr20015y, 15) 
apr20015ze <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "5Ze")
apr20015ze <- sample_n(apr20015ze, 5) 
apr2001hud <- subset(backno1, year == "2001" & month == "4" & nafo_zone == "HudsonStrait")
apr2001hud <- sample_n(apr2001hud, 39) 
apr2001 <- rbind(apr20010a, apr20010b, apr20011B, apr20011C, apr20011d, apr20011E, apr20011F,  apr20012G, apr20012H, apr20012J, apr20013k, apr20013L, apr20013M, apr20013N, apr20013o, apr20013pn, apr20013ps, apr20014r, apr20014s, apr20014t, apr20014vn, apr20014vs, apr20014w, apr20014x, apr20015y, apr20015ze, apr2001hud)
```

```{r apr 2002}
apr20020a <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "0A")
apr20020a <- sample_n(apr20020a, 25) 
apr20020b <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "0B")
apr20020b <- sample_n(apr20020b, 103) 
apr20021B <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "1B")
apr20021B <- sample_n(apr20021B, 2)
apr20021C <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "1C")
apr20021C <- sample_n(apr20021C, 9) 
apr20021d <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "1D")
apr20021d <- sample_n(apr20021d, 16)
apr20021E <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "1E")
apr20021E <- sample_n(apr20021E, 22)
apr20021F <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "1F")
apr20021F <- sample_n(apr20021F, 41)
apr20022G <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "2G")
apr20022G <- sample_n(apr20022G, 72) 
apr20022H <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "2H")
apr20022H <- sample_n(apr20022H, 76) 
apr20022J <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "2J")
apr20022J <- sample_n(apr20022J, 90) 
apr20023k <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "3K")
apr20023k <- sample_n(apr20023k, 94) 
apr20023L <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "3L")
apr20023L <- sample_n(apr20023L, 79)
apr20023M <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "3M")
apr20023M <- sample_n(apr20023M, 5)
apr20023N <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "3N")
apr20023N <- sample_n(apr20023N, 29) 
apr20023o <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "3O")
apr20023o <- sample_n(apr20023o, 41) 
apr20023pn <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "3Pn")
apr20023pn <- sample_n(apr20023pn, 4) 
apr20023ps <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "3Ps")
apr20023ps <- sample_n(apr20023ps, 42) 
apr20024r <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "4R")
apr20024r <- sample_n(apr20024r, 20) 
apr20024s <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "4S")
apr20024s <- sample_n(apr20024s, 39) 
apr20024t <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "4T")
apr20024t <- sample_n(apr20024t, 54) 
apr20024vn <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "4Vn")
apr20024vn <- sample_n(apr20024vn, 11) 
apr20024vs <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "4Vs")
apr20024vs <- sample_n(apr20024vs, 42) 
apr20024w <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "4W")
apr20024w <- sample_n(apr20024w, 43) 
apr20024x <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "4X")
apr20024x <- sample_n(apr20024x, 40) 
apr20025y <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "5Y")
apr20025y <- sample_n(apr20025y, 16) 
apr20025ze <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "5Ze")
apr20025ze <- sample_n(apr20025ze, 5) 
apr2002hud <- subset(backno1, year == "2002" & month == "4" & nafo_zone == "HudsonStrait")
apr2002hud <- sample_n(apr2002hud, 42) 
apr2002 <- rbind(apr20020a, apr20020b, apr20021B, apr20021C, apr20021d, apr20021E, apr20021F,  apr20022G, apr20022H, apr20022J, apr20023k, apr20023L, apr20023M, apr20023N, apr20023o, apr20023pn, apr20023ps, apr20024r, apr20024s, apr20024t, apr20024vn, apr20024vs, apr20024w, apr20024x, apr20025y, apr20025ze, apr2002hud)
```

```{r apr 2003}
apr20030a <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "0A")
apr20030a <- sample_n(apr20030a, 24) 
apr20030b <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "0B")
apr20030b <- sample_n(apr20030b, 97) 
apr20031B <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "1B")
apr20031B <- sample_n(apr20031B, 2)
apr20031C <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "1C")
apr20031C <- sample_n(apr20031C, 8) 
apr20031d <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "1D")
apr20031d <- sample_n(apr20031d, 16)
apr20031E <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "1E")
apr20031E <- sample_n(apr20031E, 20)
apr20031F <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "1F")
apr20031F <- sample_n(apr20031F, 38)
apr20032G <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "2G")
apr20032G <- sample_n(apr20032G, 68) 
apr20032H <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "2H")
apr20032H <- sample_n(apr20032H, 72) 
apr20032J <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "2J")
apr20032J <- sample_n(apr20032J, 85) 
apr20033k <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "3K")
apr20033k <- sample_n(apr20033k, 88) 
apr20033L <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "3L")
apr20033L <- sample_n(apr20033L, 74)
apr20033M <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "3M")
apr20033M <- sample_n(apr20033M, 5)
apr20033N <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "3N")
apr20033N <- sample_n(apr20033N, 27) 
apr20033o <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "3O")
apr20033o <- sample_n(apr20033o, 39) 
apr20033pn <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "3Pn")
apr20033pn <- sample_n(apr20033pn, 4) 
apr20033ps <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "3Ps")
apr20033ps <- sample_n(apr20033ps, 40) 
apr20034r <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "4R")
apr20034r <- sample_n(apr20034r, 19) 
apr20034s <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "4S")
apr20034s <- sample_n(apr20034s, 37) 
apr20034t <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "4T")
apr20034t <- sample_n(apr20034t, 51) 
apr20034vn <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "4Vn")
apr20034vn <- sample_n(apr20034vn, 11) 
apr20034vs <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "4Vs")
apr20034vs <- sample_n(apr20034vs, 40) 
apr20034w <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "4W")
apr20034w <- sample_n(apr20034w, 41) 
apr20034x <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "4X")
apr20034x <- sample_n(apr20034x, 38) 
apr20035y <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "5Y")
apr20035y <- sample_n(apr20035y, 15) 
apr20035ze <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "5Ze")
apr20035ze <- sample_n(apr20035ze, 5) 
apr2003hud <- subset(backno1, year == "2003" & month == "4" & nafo_zone == "HudsonStrait")
apr2003hud <- sample_n(apr2003hud, 40) 
apr2003 <- rbind(apr20030a, apr20030b, apr20031B, apr20031C, apr20031d, apr20031E, apr20031F,  apr20032G, apr20032H, apr20032J, apr20033k, apr20033L, apr20033M, apr20033N, apr20033o, apr20033pn, apr20033ps, apr20034r, apr20034s, apr20034t, apr20034vn, apr20034vs, apr20034w, apr20034x, apr20035y, apr20035ze, apr2003hud)
```

```{r apr 2004}
apr20040a <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "0A")
apr20040a <- sample_n(apr20040a, 17) 
apr20040b <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "0B")
apr20040b <- sample_n(apr20040b, 71) 
apr20041B <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "1B")
apr20041B <- sample_n(apr20041B, 1)
apr20041C <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "1C")
apr20041C <- sample_n(apr20041C, 6) 
apr20041d <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "1D")
apr20041d <- sample_n(apr20041d, 11)
apr20041E <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "1E")
apr20041E <- sample_n(apr20041E, 15)
apr20041F <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "1F")
apr20041F <- sample_n(apr20041F, 28)
apr20042G <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "2G")
apr20042G <- sample_n(apr20042G, 50) 
apr20042H <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "2H")
apr20042H <- sample_n(apr20042H, 53) 
apr20042J <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "2J")
apr20042J <- sample_n(apr20042J, 62) 
apr20043k <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "3K")
apr20043k <- sample_n(apr20043k, 64) 
apr20043L <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "3L")
apr20043L <- sample_n(apr20043L, 54)
apr20043M <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "3M")
apr20043M <- sample_n(apr20043M, 3)
apr20043N <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "3N")
apr20043N <- sample_n(apr20043N, 20) 
apr20043o <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "3O")
apr20043o <- sample_n(apr20043o, 28) 
apr20043pn <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "3Pn")
apr20043pn <- sample_n(apr20043pn, 3) 
apr20043ps <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "3Ps")
apr20043ps <- sample_n(apr20043ps, 29) 
apr20044r <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "4R")
apr20044r <- sample_n(apr20044r, 14) 
apr20044s <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "4S")
apr20044s <- sample_n(apr20044s, 27) 
apr20044t <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "4T")
apr20044t <- sample_n(apr20044t, 37) 
apr20044vn <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "4Vn")
apr20044vn <- sample_n(apr20044vn, 8) 
apr20044vs <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "4Vs")
apr20044vs <- sample_n(apr20044vs, 29) 
apr20044w <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "4W")
apr20044w <- sample_n(apr20044w, 30) 
apr20044x <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "4X")
apr20044x <- sample_n(apr20044x, 28) 
apr20045y <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "5Y")
apr20045y <- sample_n(apr20045y, 11) 
apr20045ze <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "5Ze")
apr20045ze <- sample_n(apr20045ze, 4) 
apr2004hud <- subset(backno1, year == "2004" & month == "4" & nafo_zone == "HudsonStrait")
apr2004hud <- sample_n(apr2004hud, 29) 
apr2004 <- rbind(apr20040a, apr20040b, apr20041B, apr20041C, apr20041d, apr20041E, apr20041F,  apr20042G, apr20042H, apr20042J, apr20043k, apr20043L, apr20043M, apr20043N, apr20043o, apr20043pn, apr20043ps, apr20044r, apr20044s, apr20044t, apr20044vn, apr20044vs, apr20044w, apr20044x, apr20045y, apr20045ze, apr2004hud)
```

```{r apr 2005}
apr20050a <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "0A")
apr20050a <- sample_n(apr20050a, 6) 
apr20050b <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "0B")
apr20050b <- sample_n(apr20050b, 30) 
apr20051B <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "1B")
apr20051B <- sample_n(apr20051B, 1)
apr20051C <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "1C")
apr20051C <- sample_n(apr20051C, 3) 
apr20051d <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "1D")
apr20051d <- sample_n(apr20051d, 5)
apr20051E <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "1E")
apr20051E <- sample_n(apr20051E, 6)
apr20051F <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "1F")
apr20051F <- sample_n(apr20051F, 12)
apr20052G <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "2G")
apr20052G <- sample_n(apr20052G, 21) 
apr20052H <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "2H")
apr20052H <- sample_n(apr20052H, 23) 
apr20052J <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "2J")
apr20052J <- sample_n(apr20052J, 27) 
apr20053k <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "3K")
apr20053k <- sample_n(apr20053k, 28) 
apr20053L <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "3L")
apr20053L <- sample_n(apr20053L, 23)
apr20053M <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "3M")
apr20053M <- sample_n(apr20053M, 2)
apr20053N <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "3N")
apr20053N <- sample_n(apr20053N, 8) 
apr20053o <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "3O")
apr20053o <- sample_n(apr20053o, 12) 
apr20053pn <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "3Pn")
apr20053pn <- sample_n(apr20053pn, 1) 
apr20053ps <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "3Ps")
apr20053ps <- sample_n(apr20053ps, 13) 
apr20054r <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "4R")
apr20054r <- sample_n(apr20054r, 6) 
apr20054s <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "4S")
apr20054s <- sample_n(apr20054s, 12) 
apr20054t <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "4T")
apr20054t <- sample_n(apr20054t, 16) 
apr20054vn <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "4Vn")
apr20054vn <- sample_n(apr20054vn, 3) 
apr20054vs <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "4Vs")
apr20054vs <- sample_n(apr20054vs, 13) 
apr20054w <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "4W")
apr20054w <- sample_n(apr20054w, 13) 
apr20054x <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "4X")
apr20054x <- sample_n(apr20054x, 12) 
apr20055y <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "5Y")
apr20055y <- sample_n(apr20055y, 5) 
apr20055ze <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "5Ze")
apr20055ze <- sample_n(apr20055ze, 2) 
apr2005hud <- subset(backno1, year == "2005" & month == "4" & nafo_zone == "HudsonStrait")
apr2005hud <- sample_n(apr2005hud, 13) 
apr2005 <- rbind(apr20050a, apr20050b, apr20051B, apr20051C, apr20051d, apr20051E, apr20051F,  apr20052G, apr20052H, apr20052J, apr20053k, apr20053L, apr20053M, apr20053N, apr20053o, apr20053pn, apr20053ps, apr20054r, apr20054s, apr20054t, apr20054vn, apr20054vs, apr20054w, apr20054x, apr20055y, apr20055ze, apr2005hud)
```

```{r apr 2006}
apr20060a <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "0A")
apr20060a <- sample_n(apr20060a, 7) 
apr20060b <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "0B")
apr20060b <- sample_n(apr20060b, 29) 
apr20061B <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "1B")
apr20061B <- sample_n(apr20061B, 1)
apr20061C <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "1C")
apr20061C <- sample_n(apr20061C, 3) 
apr20061d <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "1D")
apr20061d <- sample_n(apr20061d, 5)
apr20061E <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "1E")
apr20061E <- sample_n(apr20061E, 6)
apr20061F <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "1F")
apr20061F <- sample_n(apr20061F, 11)
apr20062G <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "2G")
apr20062G <- sample_n(apr20062G, 20) 
apr20062H <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "2H")
apr20062H <- sample_n(apr20062H, 22) 
apr20062J <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "2J")
apr20062J <- sample_n(apr20062J, 26) 
apr20063k <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "3K")
apr20063k <- sample_n(apr20063k, 27) 
apr20063L <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "3L")
apr20063L <- sample_n(apr20063L, 22)
apr20063M <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "3M")
apr20063M <- sample_n(apr20063M, 1)
apr20063N <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "3N")
apr20063N <- sample_n(apr20063N, 8) 
apr20063o <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "3O")
apr20063o <- sample_n(apr20063o, 12) 
apr20063pn <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "3Pn")
apr20063pn <- sample_n(apr20063pn, 1) 
apr20063ps <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "3Ps")
apr20063ps <- sample_n(apr20063ps, 12) 
apr20064r <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "4R")
apr20064r <- sample_n(apr20064r, 6) 
apr20064s <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "4S")
apr20064s <- sample_n(apr20064s, 11) 
apr20064t <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "4T")
apr20064t <- sample_n(apr20064t, 15) 
apr20064vn <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "4Vn")
apr20064vn <- sample_n(apr20064vn, 3) 
apr20064vs <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "4Vs")
apr20064vs <- sample_n(apr20064vs, 12) 
apr20064w <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "4W")
apr20064w <- sample_n(apr20064w, 12) 
apr20064x <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "4X")
apr20064x <- sample_n(apr20064x, 11) 
apr20065y <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "5Y")
apr20065y <- sample_n(apr20065y, 5) 
apr20065ze <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "5Ze")
apr20065ze <- sample_n(apr20065ze, 1) 
apr2006hud <- subset(backno1, year == "2006" & month == "4" & nafo_zone == "HudsonStrait")
apr2006hud <- sample_n(apr2006hud, 12) 
apr2006 <- rbind(apr20060a, apr20060b, apr20061B, apr20061C, apr20061d, apr20061E, apr20061F,  apr20062G, apr20062H, apr20062J, apr20063k, apr20063L, apr20063M, apr20063N, apr20063o, apr20063pn, apr20063ps, apr20064r, apr20064s, apr20064t, apr20064vn, apr20064vs, apr20064w, apr20064x, apr20065y, apr20065ze, apr2006hud)
```


```{r apr 2007}
apr20070a <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "0A")
apr20070a <- sample_n(apr20070a, 18) 
apr20070b <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "0B")
apr20070b <- sample_n(apr20070b, 72) 
apr20071B <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "1B")
apr20071B <- sample_n(apr20071B, 1)
apr20071C <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "1C")
apr20071C <- sample_n(apr20071C, 6) 
apr20071d <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "1D")
apr20071d <- sample_n(apr20071d, 12)
apr20071E <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "1E")
apr20071E <- sample_n(apr20071E, 15)
apr20071F <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "1F")
apr20071F <- sample_n(apr20071F, 28)
apr20072G <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "2G")
apr20072G <- sample_n(apr20072G, 51) 
apr20072H <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "2H")
apr20072H <- sample_n(apr20072H, 54) 
apr20072J <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "2J")
apr20072J <- sample_n(apr20072J, 63) 
apr20073k <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "3K")
apr20073k <- sample_n(apr20073k, 66) 
apr20073L <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "3L")
apr20073L <- sample_n(apr20073L, 55)
apr20073M <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "3M")
apr20073M <- sample_n(apr20073M, 3)
apr20073N <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "3N")
apr20073N <- sample_n(apr20073N, 20) 
apr20073o <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "3O")
apr20073o <- sample_n(apr20073o, 29) 
apr20073pn <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "3Pn")
apr20073pn <- sample_n(apr20073pn, 3) 
apr20073ps <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "3Ps")
apr20073ps <- sample_n(apr20073ps, 30) 
apr20074r <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "4R")
apr20074r <- sample_n(apr20074r, 14) 
apr20074s <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "4S")
apr20074s <- sample_n(apr20074s, 27) 
apr20074t <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "4T")
apr20074t <- sample_n(apr20074t, 38) 
apr20074vn <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "4Vn")
apr20074vn <- sample_n(apr20074vn, 8) 
apr20074vs <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "4Vs")
apr20074vs <- sample_n(apr20074vs, 30) 
apr20074w <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "4W")
apr20074w <- sample_n(apr20074w, 30) 
apr20074x <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "4X")
apr20074x <- sample_n(apr20074x, 28) 
apr20075y <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "5Y")
apr20075y <- sample_n(apr20075y, 11) 
apr20075ze <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "5Ze")
apr20075ze <- sample_n(apr20075ze, 4) 
apr2007hud <- subset(backno1, year == "2007" & month == "4" & nafo_zone == "HudsonStrait")
apr2007hud <- sample_n(apr2007hud, 30) 
apr2007 <- rbind(apr20070a, apr20070b, apr20071B, apr20071C, apr20071d, apr20071E, apr20071F,  apr20072G, apr20072H, apr20072J, apr20073k, apr20073L, apr20073M, apr20073N, apr20073o, apr20073pn, apr20073ps, apr20074r, apr20074s, apr20074t, apr20074vn, apr20074vs, apr20074w, apr20074x, apr20075y, apr20075ze, apr2007hud)
```


```{r apr 2008}
apr20080a <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "0A")
apr20080a <- sample_n(apr20080a, 12) 
apr20080b <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "0B")
apr20080b <- sample_n(apr20080b, 50) 
apr20081B <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "1B")
apr20081B <- sample_n(apr20081B, 1)
apr20081C <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "1C")
apr20081C <- sample_n(apr20081C, 4) 
apr20081d <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "1D")
apr20081d <- sample_n(apr20081d, 8)
apr20081E <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "1E")
apr20081E <- sample_n(apr20081E, 10)
apr20081F <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "1F")
apr20081F <- sample_n(apr20081F, 20)
apr20082G <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "2G")
apr20082G <- sample_n(apr20082G, 35) 
apr20082H <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "2H")
apr20082H <- sample_n(apr20082H, 37) 
apr20082J <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "2J")
apr20082J <- sample_n(apr20082J, 44) 
apr20083k <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "3K")
apr20083k <- sample_n(apr20083k, 45) 
apr20083L <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "3L")
apr20083L <- sample_n(apr20083L, 38)
apr20083M <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "3M")
apr20083M <- sample_n(apr20083M, 2)
apr20083N <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "3N")
apr20083N <- sample_n(apr20083N, 14) 
apr20083o <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "3O")
apr20083o <- sample_n(apr20083o, 20) 
apr20083pn <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "3Pn")
apr20083pn <- sample_n(apr20083pn, 2) 
apr20083ps <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "3Ps")
apr20083ps <- sample_n(apr20083ps, 21) 
apr20084r <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "4R")
apr20084r <- sample_n(apr20084r, 10) 
apr20084s <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "4S")
apr20084s <- sample_n(apr20084s, 19) 
apr20084t <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "4T")
apr20084t <- sample_n(apr20084t, 26) 
apr20084vn <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "4Vn")
apr20084vn <- sample_n(apr20084vn, 5) 
apr20084vs <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "4Vs")
apr20084vs <- sample_n(apr20084vs, 21) 
apr20084w <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "4W")
apr20084w <- sample_n(apr20084w, 21) 
apr20084x <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "4X")
apr20084x <- sample_n(apr20084x, 19) 
apr20085y <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "5Y")
apr20085y <- sample_n(apr20085y, 8) 
apr20085ze <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "5Ze")
apr20085ze <- sample_n(apr20085ze, 3) 
apr2008hud <- subset(backno1, year == "2008" & month == "4" & nafo_zone == "HudsonStrait")
apr2008hud <- sample_n(apr2008hud, 21) 
apr2008 <- rbind(apr20080a, apr20080b, apr20081B, apr20081C, apr20081d, apr20081E, apr20081F,  apr20082G, apr20082H, apr20082J, apr20083k, apr20083L, apr20083M, apr20083N, apr20083o, apr20083pn, apr20083ps, apr20084r, apr20084s, apr20084t, apr20084vn, apr20084vs, apr20084w, apr20084x, apr20085y, apr20085ze, apr2008hud)
```

```{r apr 2009}
apr20090a <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "0A")
apr20090a <- sample_n(apr20090a, 22) 
apr20090b <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "0B")
apr20090b <- sample_n(apr20090b, 92) 
apr20091B <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "1B")
apr20091B <- sample_n(apr20091B, 2)
apr20091C <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "1C")
apr20091C <- sample_n(apr20091C, 8) 
apr20091d <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "1D")
apr20091d <- sample_n(apr20091d, 15)
apr20091E <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "1E")
apr20091E <- sample_n(apr20091E, 19)
apr20091F <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "1F")
apr20091F <- sample_n(apr20091F, 36)
apr20092G <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "2G")
apr20092G <- sample_n(apr20092G, 64) 
apr20092H <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "2H")
apr20092H <- sample_n(apr20092H, 68) 
apr20092J <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "2J")
apr20092J <- sample_n(apr20092J, 80) 
apr20093k <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "3K")
apr20093k <- sample_n(apr20093k, 83) 
apr20093L <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "3L")
apr20093L <- sample_n(apr20093L, 70)
apr20093M <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "3M")
apr20093M <- sample_n(apr20093M, 4)
apr20093N <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "3N")
apr20093N <- sample_n(apr20093N, 25) 
apr20093o <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "3O")
apr20093o <- sample_n(apr20093o, 37) 
apr20093pn <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "3Pn")
apr20093pn <- sample_n(apr20093pn, 3) 
apr20093ps <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "3Ps")
apr20093ps <- sample_n(apr20093ps, 38) 
apr20094r <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "4R")
apr20094r <- sample_n(apr20094r, 18) 
apr20094s <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "4S")
apr20094s <- sample_n(apr20094s, 35) 
apr20094t <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "4T")
apr20094t <- sample_n(apr20094t, 48) 
apr20094vn <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "4Vn")
apr20094vn <- sample_n(apr20094vn, 10) 
apr20094vs <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "4Vs")
apr20094vs <- sample_n(apr20094vs, 38) 
apr20094w <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "4W")
apr20094w <- sample_n(apr20094w, 38) 
apr20094x <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "4X")
apr20094x <- sample_n(apr20094x, 36) 
apr20095y <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "5Y")
apr20095y <- sample_n(apr20095y, 14) 
apr20095ze <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "5Ze")
apr20095ze <- sample_n(apr20095ze, 5) 
apr2009hud <- subset(backno1, year == "2009" & month == "4" & nafo_zone == "HudsonStrait")
apr2009hud <- sample_n(apr2009hud, 38) 
apr2009 <- rbind(apr20090a, apr20090b, apr20091B, apr20091C, apr20091d, apr20091E, apr20091F,  apr20092G, apr20092H, apr20092J, apr20093k, apr20093L, apr20093M, apr20093N, apr20093o, apr20093pn, apr20093ps, apr20094r, apr20094s, apr20094t, apr20094vn, apr20094vs, apr20094w, apr20094x, apr20095y, apr20095ze, apr2009hud)
```

```{r apr 2010}
apr20100a <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "0A")
apr20100a <- sample_n(apr20100a, 13) 
apr20100b <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "0B")
apr20100b <- sample_n(apr20100b, 54) 
apr20101B <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "1B")
apr20101B <- sample_n(apr20101B, 1)
apr20101C <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "1C")
apr20101C <- sample_n(apr20101C, 5) 
apr20101d <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "1D")
apr20101d <- sample_n(apr20101d, 9)
apr20101E <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "1E")
apr20101E <- sample_n(apr20101E, 11)
apr20101F <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "1F")
apr20101F <- sample_n(apr20101F, 21)
apr20102G <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "2G")
apr20102G <- sample_n(apr20102G, 38) 
apr20102H <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "2H")
apr20102H <- sample_n(apr20102H, 40) 
apr20102J <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "2J")
apr20102J <- sample_n(apr20102J, 48) 
apr20103k <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "3K")
apr20103k <- sample_n(apr20103k, 49) 
apr20103L <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "3L")
apr20103L <- sample_n(apr20103L, 41)
apr20103M <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "3M")
apr20103M <- sample_n(apr20103M, 3)
apr20103N <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "3N")
apr20103N <- sample_n(apr20103N, 15) 
apr20103o <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "3O")
apr20103o <- sample_n(apr20103o, 22) 
apr20103pn <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "3Pn")
apr20103pn <- sample_n(apr20103pn, 2) 
apr20103ps <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "3Ps")
apr20103ps <- sample_n(apr20103ps, 22) 
apr20104r <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "4R")
apr20104r <- sample_n(apr20104r, 11) 
apr20104s <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "4S")
apr20104s <- sample_n(apr20104s, 21) 
apr20104t <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "4T")
apr20104t <- sample_n(apr20104t, 29) 
apr20104vn <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "4Vn")
apr20104vn <- sample_n(apr20104vn, 6) 
apr20104vs <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "4Vs")
apr20104vs <- sample_n(apr20104vs, 22) 
apr20104w <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "4W")
apr20104w <- sample_n(apr20104w, 23) 
apr20104x <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "4X")
apr20104x <- sample_n(apr20104x, 21) 
apr20105y <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "5Y")
apr20105y <- sample_n(apr20105y, 8) 
apr20105ze <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "5Ze")
apr20105ze <- sample_n(apr20105ze, 3) 
apr2010hud <- subset(backno1, year == "2010" & month == "4" & nafo_zone == "HudsonStrait")
apr2010hud <- sample_n(apr2010hud, 22) 
apr2010 <- rbind(apr20100a, apr20100b, apr20101B, apr20101C, apr20101d, apr20101E, apr20101F,  apr20102G, apr20102H, apr20102J, apr20103k, apr20103L, apr20103M, apr20103N, apr20103o, apr20103pn, apr20103ps, apr20104r, apr20104s, apr20104t, apr20104vn, apr20104vs, apr20104w, apr20104x, apr20105y, apr20105ze, apr2010hud)
```

```{r apr 2011}
apr20110a <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "0A")
apr20110a <- sample_n(apr20110a, 18) 
apr20110b <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "0B")
apr20110b <- sample_n(apr20110b, 75) 
apr20111B <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "1B")
apr20111B <- sample_n(apr20111B, 1)
apr20111C <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "1C")
apr20111C <- sample_n(apr20111C, 6) 
apr20111d <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "1D")
apr20111d <- sample_n(apr20111d, 12)
apr20111E <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "1E")
apr20111E <- sample_n(apr20111E, 16)
apr20111F <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "1F")
apr20111F <- sample_n(apr20111F, 30)
apr20112G <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "2G")
apr20112G <- sample_n(apr20112G, 53) 
apr20112H <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "2H")
apr20112H <- sample_n(apr20112H, 56) 
apr20112J <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "2J")
apr20112J <- sample_n(apr20112J, 66) 
apr20113k <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "3K")
apr20113k <- sample_n(apr20113k, 68) 
apr20113L <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "3L")
apr20113L <- sample_n(apr20113L, 57)
apr20113M <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "3M")
apr20113M <- sample_n(apr20113M, 4)
apr20113N <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "3N")
apr20113N <- sample_n(apr20113N, 21) 
apr20113o <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "3O")
apr20113o <- sample_n(apr20113o, 30) 
apr20113pn <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "3Pn")
apr20113pn <- sample_n(apr20113pn, 3) 
apr20113ps <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "3Ps")
apr20113ps <- sample_n(apr20113ps, 31) 
apr20114r <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "4R")
apr20114r <- sample_n(apr20114r, 15) 
apr20114s <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "4S")
apr20114s <- sample_n(apr20114s, 28) 
apr20114t <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "4T")
apr20114t <- sample_n(apr20114t, 39) 
apr20114vn <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "4Vn")
apr20114vn <- sample_n(apr20114vn, 8) 
apr20114vs <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "4Vs")
apr20114vs <- sample_n(apr20114vs, 31) 
apr20114w <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "4W")
apr20114w <- sample_n(apr20114w, 31) 
apr20114x <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "4X")
apr20114x <- sample_n(apr20114x, 29) 
apr20115y <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "5Y")
apr20115y <- sample_n(apr20115y, 12) 
apr20115ze <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "5Ze")
apr20115ze <- sample_n(apr20115ze, 4) 
apr2011hud <- subset(backno1, year == "2011" & month == "4" & nafo_zone == "HudsonStrait")
apr2011hud <- sample_n(apr2011hud, 31) 
apr2011 <- rbind(apr20110a, apr20110b, apr20111B, apr20111C, apr20111d, apr20111E, apr20111F,  apr20112G, apr20112H, apr20112J, apr20113k, apr20113L, apr20113M, apr20113N, apr20113o, apr20113pn, apr20113ps, apr20114r, apr20114s, apr20114t, apr20114vn, apr20114vs, apr20114w, apr20114x, apr20115y, apr20115ze, apr2011hud)
```

```{r merge all ar}
apr <- rbind(apr1998, apr1999, apr2000, apr2001, apr2002, apr2003, apr2004, apr2005, apr2006, apr2007, apr2008, apr2009, apr2010, apr2011)
```



```{r may 1998}
may19980a <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "0A")
may19980a <- sample_n(may19980a, 19) 
may19980b <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "0B")
may19980b <- sample_n(may19980b, 77) 
may19981B <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "1B")
may19981B <- sample_n(may19981B, 1)
may19981C <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "1C")
may19981C <- sample_n(may19981C, 7) 
may19981d <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "1D")
may19981d <- sample_n(may19981d, 12)
may19981E <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "1E")
may19981E <- sample_n(may19981E, 16)
may19981F <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "1F")
may19981F <- sample_n(may19981F, 30)
may19982G <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "2G")
may19982G <- sample_n(may19982G, 54) 
may19982H <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "2H")
may19982H <- sample_n(may19982H, 57) 
may19982J <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "2J")
may19982J <- sample_n(may19982J, 67) 
may19983k <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "3K")
may19983k <- sample_n(may19983k, 70) 
may19983L <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "3L")
may19983L <- sample_n(may19983L, 58)
may19983M <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "3M")
may19983M <- sample_n(may19983M, 4)
may19983N <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "3N")
may19983N <- sample_n(may19983N, 21) 
may19983o <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "3O")
may19983o <- sample_n(may19983o, 31) 
may19983pn <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "3Pn")
may19983pn <- sample_n(may19983pn, 3) 
may19983ps <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "3Ps")
may19983ps <- sample_n(may19983ps, 31) 
may19984r <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "4R")
may19984r <- sample_n(may19984r, 15) 
may19984s <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "4S")
may19984s <- sample_n(may19984s, 29) 
may19984t <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "4T")
may19984t <- sample_n(may19984t, 40) 
may19984vn <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "4Vn")
may19984vn <- sample_n(may19984vn, 8) 
may19984vs <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "4Vs")
may19984vs <- sample_n(may19984vs, 31) 
may19984w <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "4W")
may19984w <- sample_n(may19984w, 32) 
may19984x <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "4X")
may19984x <- sample_n(may19984x, 30) 
may19985y <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "5Y")
may19985y <- sample_n(may19985y, 12) 
may19985ze <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "5Ze")
may19985ze <- sample_n(may19985ze, 4) 
may1998hud <- subset(backno1, year == "1998" & month == "5" & nafo_zone == "HudsonStrait")
may1998hud <- sample_n(may1998hud, 31) 
may1998 <- rbind(may19980a, may19980b, may19981B, may19981C, may19981d, may19981E, may19981F,  may19982G, may19982H, may19982J, may19983k, may19983L, may19983M, may19983N, may19983o, may19983pn, may19983ps, may19984r, may19984s, may19984t, may19984vn, may19984vs, may19984w, may19984x, may19985y, may19985ze, may1998hud)
```

```{r may 1999}
may19990a <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "0A")
may19990a <- sample_n(may19990a, 14) 
may19990b <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "0B")
may19990b <- sample_n(may19990b, 58) 
may19991B <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "1B")
may19991B <- sample_n(may19991B, 1)
may19991C <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "1C")
may19991C <- sample_n(may19991C, 5) 
may19991d <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "1D")
may19991d <- sample_n(may19991d, 9)
may19991E <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "1E")
may19991E <- sample_n(may19991E, 12)
may19991F <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "1F")
may19991F <- sample_n(may19991F, 23)
may19992G <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "2G")
may19992G <- sample_n(may19992G, 40) 
may19992H <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "2H")
may19992H <- sample_n(may19992H, 43) 
may19992J <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "2J")
may19992J <- sample_n(may19992J, 50) 
may19993k <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "3K")
may19993k <- sample_n(may19993k, 52) 
may19993L <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "3L")
may19993L <- sample_n(may19993L, 44)
may19993M <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "3M")
may19993M <- sample_n(may19993M, 3)
may19993N <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "3N")
may19993N <- sample_n(may19993N, 16) 
may19993o <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "3O")
may19993o <- sample_n(may19993o, 23) 
may19993pn <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "3Pn")
may19993pn <- sample_n(may19993pn, 2) 
may19993ps <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "3Ps")
may19993ps <- sample_n(may19993ps, 24) 
may19994r <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "4R")
may19994r <- sample_n(may19994r, 11) 
may19994s <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "4S")
may19994s <- sample_n(may19994s, 22) 
may19994t <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "4T")
may19994t <- sample_n(may19994t, 30) 
may19994vn <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "4Vn")
may19994vn <- sample_n(may19994vn, 6) 
may19994vs <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "4Vs")
may19994vs <- sample_n(may19994vs, 24) 
may19994w <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "4W")
may19994w <- sample_n(may19994w, 24) 
may19994x <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "4X")
may19994x <- sample_n(may19994x, 23) 
may19995y <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "5Y")
may19995y <- sample_n(may19995y, 9) 
may19995ze <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "5Ze")
may19995ze <- sample_n(may19995ze, 3) 
may1999hud <- subset(backno1, year == "1999" & month == "5" & nafo_zone == "HudsonStrait")
may1999hud <- sample_n(may1999hud, 24) 
may1999 <- rbind(may19990a, may19990b, may19991B, may19991C, may19991d, may19991E, may19991F,  may19992G, may19992H, may19992J, may19993k, may19993L, may19993M, may19993N, may19993o, may19993pn, may19993ps, may19994r, may19994s, may19994t, may19994vn, may19994vs, may19994w, may19994x, may19995y, may19995ze, may1999hud)
```


```{r may 2000}
may20000a <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "0A")
may20000a <- sample_n(may20000a, 13) 
may20000b <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "0B")
may20000b <- sample_n(may20000b, 52) 
may20001B <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "1B")
may20001B <- sample_n(may20001B, 1)
may20001C <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "1C")
may20001C <- sample_n(may20001C, 4) 
may20001d <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "1D")
may20001d <- sample_n(may20001d, 8)
may20001E <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "1E")
may20001E <- sample_n(may20001E, 11)
may20001F <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "1F")
may20001F <- sample_n(may20001F, 20)
may20002G <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "2G")
may20002G <- sample_n(may20002G, 37) 
may20002H <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "2H")
may20002H <- sample_n(may20002H, 39) 
may20002J <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "2J")
may20002J <- sample_n(may20002J, 46) 
may20003k <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "3K")
may20003k <- sample_n(may20003k, 47) 
may20003L <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "3L")
may20003L <- sample_n(may20003L, 40)
may20003M <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "3M")
may20003M <- sample_n(may20003M, 3)
may20003N <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "3N")
may20003N <- sample_n(may20003N, 14) 
may20003o <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "3O")
may20003o <- sample_n(may20003o, 21) 
may20003pn <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "3Pn")
may20003pn <- sample_n(may20003pn, 2) 
may20003ps <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "3Ps")
may20003ps <- sample_n(may20003ps, 21) 
may20004r <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "4R")
may20004r <- sample_n(may20004r, 10) 
may20004s <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "4S")
may20004s <- sample_n(may20004s, 20) 
may20004t <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "4T")
may20004t <- sample_n(may20004t, 27) 
may20004vn <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "4Vn")
may20004vn <- sample_n(may20004vn, 6) 
may20004vs <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "4Vs")
may20004vs <- sample_n(may20004vs, 21) 
may20004w <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "4W")
may20004w <- sample_n(may20004w, 22) 
may20004x <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "4X")
may20004x <- sample_n(may20004x, 20) 
may20005y <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "5Y")
may20005y <- sample_n(may20005y, 8) 
may20005ze <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "5Ze")
may20005ze <- sample_n(may20005ze, 3) 
may2000hud <- subset(backno1, year == "2000" & month == "5" & nafo_zone == "HudsonStrait")
may2000hud <- sample_n(may2000hud, 21) 
may2000 <- rbind(may20000a, may20000b, may20001B, may20001C, may20001d, may20001E, may20001F,  may20002G, may20002H, may20002J, may20003k, may20003L, may20003M, may20003N, may20003o, may20003pn, may20003ps, may20004r, may20004s, may20004t, may20004vn, may20004vs, may20004w, may20004x, may20005y, may20005ze, may2000hud)
```


```{r may 2001}
may20010a <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "0A")
may20010a <- sample_n(may20010a, 18) 
may20010b <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "0B")
may20010b <- sample_n(may20010b, 73) 
may20011B <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "1B")
may20011B <- sample_n(may20011B, 1)
may20011C <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "1C")
may20011C <- sample_n(may20011C, 6) 
may20011d <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "1D")
may20011d <- sample_n(may20011d, 12)
may20011E <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "1E")
may20011E <- sample_n(may20011E, 15)
may20011F <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "1F")
may20011F <- sample_n(may20011F, 28)
may20012G <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "2G")
may20012G <- sample_n(may20012G, 51) 
may20012H <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "2H")
may20012H <- sample_n(may20012H, 54) 
may20012J <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "2J")
may20012J <- sample_n(may20012J, 64) 
may20013k <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "3K")
may20013k <- sample_n(may20013k, 66) 
may20013L <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "3L")
may20013L <- sample_n(may20013L, 56)
may20013M <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "3M")
may20013M <- sample_n(may20013M, 4)
may20013N <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "3N")
may20013N <- sample_n(may20013N, 20) 
may20013o <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "3O")
may20013o <- sample_n(may20013o, 29) 
may20013pn <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "3Pn")
may20013pn <- sample_n(may20013pn, 3) 
may20013ps <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "3Ps")
may20013ps <- sample_n(may20013ps, 30) 
may20014r <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "4R")
may20014r <- sample_n(may20014r, 14) 
may20014s <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "4S")
may20014s <- sample_n(may20014s, 28) 
may20014t <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "4T")
may20014t <- sample_n(may20014t, 38) 
may20014vn <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "4Vn")
may20014vn <- sample_n(may20014vn, 8) 
may20014vs <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "4Vs")
may20014vs <- sample_n(may20014vs, 30) 
may20014w <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "4W")
may20014w <- sample_n(may20014w, 30) 
may20014x <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "4X")
may20014x <- sample_n(may20014x, 28) 
may20015y <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "5Y")
may20015y <- sample_n(may20015y, 11) 
may20015ze <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "5Ze")
may20015ze <- sample_n(may20015ze, 4) 
may2001hud <- subset(backno1, year == "2001" & month == "5" & nafo_zone == "HudsonStrait")
may2001hud <- sample_n(may2001hud, 30) 
may2001 <- rbind(may20010a, may20010b, may20011B, may20011C, may20011d, may20011E, may20011F,  may20012G, may20012H, may20012J, may20013k, may20013L, may20013M, may20013N, may20013o, may20013pn, may20013ps, may20014r, may20014s, may20014t, may20014vn, may20014vs, may20014w, may20014x, may20015y, may20015ze, may2001hud)
```

```{r may 2002}
may20020a <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "0A")
may20020a <- sample_n(may20020a, 18) 
may20020b <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "0B")
may20020b <- sample_n(may20020b, 74) 
may20021B <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "1B")
may20021B <- sample_n(may20021B, 1)
may20021C <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "1C")
may20021C <- sample_n(may20021C, 6) 
may20021d <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "1D")
may20021d <- sample_n(may20021d, 12)
may20021E <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "1E")
may20021E <- sample_n(may20021E, 15)
may20021F <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "1F")
may20021F <- sample_n(may20021F, 29)
may20022G <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "2G")
may20022G <- sample_n(may20022G, 52) 
may20022H <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "2H")
may20022H <- sample_n(may20022H, 55) 
may20022J <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "2J")
may20022J <- sample_n(may20022J, 65) 
may20023k <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "3K")
may20023k <- sample_n(may20023k, 67) 
may20023L <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "3L")
may20023L <- sample_n(may20023L, 56)
may20023M <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "3M")
may20023M <- sample_n(may20023M, 4)
may20023N <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "3N")
may20023N <- sample_n(may20023N, 20) 
may20023o <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "3O")
may20023o <- sample_n(may20023o, 29) 
may20023pn <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "3Pn")
may20023pn <- sample_n(may20023pn, 3) 
may20023ps <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "3Ps")
may20023ps <- sample_n(may20023ps, 30) 
may20024r <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "4R")
may20024r <- sample_n(may20024r, 15) 
may20024s <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "4S")
may20024s <- sample_n(may20024s, 28) 
may20024t <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "4T")
may20024t <- sample_n(may20024t, 39) 
may20024vn <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "4Vn")
may20024vn <- sample_n(may20024vn, 8) 
may20024vs <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "4Vs")
may20024vs <- sample_n(may20024vs, 30) 
may20024w <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "4W")
may20024w <- sample_n(may20024w, 31) 
may20024x <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "4X")
may20024x <- sample_n(may20024x, 29) 
may20025y <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "5Y")
may20025y <- sample_n(may20025y, 11) 
may20025ze <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "5Ze")
may20025ze <- sample_n(may20025ze, 4) 
may2002hud <- subset(backno1, year == "2002" & month == "5" & nafo_zone == "HudsonStrait")
may2002hud <- sample_n(may2002hud, 30) 
may2002 <- rbind(may20020a, may20020b, may20021B, may20021C, may20021d, may20021E, may20021F,  may20022G, may20022H, may20022J, may20023k, may20023L, may20023M, may20023N, may20023o, may20023pn, may20023ps, may20024r, may20024s, may20024t, may20024vn, may20024vs, may20024w, may20024x, may20025y, may20025ze, may2002hud)
```

```{r may 2003}
may20030a <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "0A")
may20030a <- sample_n(may20030a, 21) 
may20030b <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "0B")
may20030b <- sample_n(may20030b, 86) 
may20031B <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "1B")
may20031B <- sample_n(may20031B, 2)
may20031C <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "1C")
may20031C <- sample_n(may20031C, 7) 
may20031d <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "1D")
may20031d <- sample_n(may20031d, 14)
may20031E <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "1E")
may20031E <- sample_n(may20031E, 18)
may20031F <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "1F")
may20031F <- sample_n(may20031F, 34)
may20032G <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "2G")
may20032G <- sample_n(may20032G, 60) 
may20032H <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "2H")
may20032H <- sample_n(may20032H, 64) 
may20032J <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "2J")
may20032J <- sample_n(may20032J, 76) 
may20033k <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "3K")
may20033k <- sample_n(may20033k, 78) 
may20033L <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "3L")
may20033L <- sample_n(may20033L, 66)
may20033M <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "3M")
may20033M <- sample_n(may20033M, 4)
may20033N <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "3N")
may20033N <- sample_n(may20033N, 24) 
may20033o <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "3O")
may20033o <- sample_n(may20033o, 34) 
may20033pn <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "3Pn")
may20033pn <- sample_n(may20033pn, 3) 
may20033ps <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "3Ps")
may20033ps <- sample_n(may20033ps, 36) 
may20034r <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "4R")
may20034r <- sample_n(may20034r, 17) 
may20034s <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "4S")
may20034s <- sample_n(may20034s, 33) 
may20034t <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "4T")
may20034t <- sample_n(may20034t, 45) 
may20034vn <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "4Vn")
may20034vn <- sample_n(may20034vn, 9) 
may20034vs <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "4Vs")
may20034vs <- sample_n(may20034vs, 35) 
may20034w <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "4W")
may20034w <- sample_n(may20034w, 36) 
may20034x <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "4X")
may20034x <- sample_n(may20034x, 34) 
may20035y <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "5Y")
may20035y <- sample_n(may20035y, 13) 
may20035ze <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "5Ze")
may20035ze <- sample_n(may20035ze, 4) 
may2003hud <- subset(backno1, year == "2003" & month == "5" & nafo_zone == "HudsonStrait")
may2003hud <- sample_n(may2003hud, 35) 
may2003 <- rbind(may20030a, may20030b, may20031B, may20031C, may20031d, may20031E, may20031F,  may20032G, may20032H, may20032J, may20033k, may20033L, may20033M, may20033N, may20033o, may20033pn, may20033ps, may20034r, may20034s, may20034t, may20034vn, may20034vs, may20034w, may20034x, may20035y, may20035ze, may2003hud)
```

```{r may 2004}
may20040a <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "0A")
may20040a <- sample_n(may20040a, 19) 
may20040b <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "0B")
may20040b <- sample_n(may20040b, 77) 
may20041B <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "1B")
may20041B <- sample_n(may20041B, 1)
may20041C <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "1C")
may20041C <- sample_n(may20041C, 7) 
may20041d <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "1D")
may20041d <- sample_n(may20041d, 12)
may20041E <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "1E")
may20041E <- sample_n(may20041E, 16)
may20041F <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "1F")
may20041F <- sample_n(may20041F, 30)
may20042G <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "2G")
may20042G <- sample_n(may20042G, 54) 
may20042H <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "2H")
may20042H <- sample_n(may20042H, 57) 
may20042J <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "2J")
may20042J <- sample_n(may20042J, 67) 
may20043k <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "3K")
may20043k <- sample_n(may20043k, 70) 
may20043L <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "3L")
may20043L <- sample_n(may20043L, 58)
may20043M <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "3M")
may20043M <- sample_n(may20043M, 4)
may20043N <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "3N")
may20043N <- sample_n(may20043N, 21) 
may20043o <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "3O")
may20043o <- sample_n(may20043o, 31) 
may20043pn <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "3Pn")
may20043pn <- sample_n(may20043pn, 3) 
may20043ps <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "3Ps")
may20043ps <- sample_n(may20043ps, 31) 
may20044r <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "4R")
may20044r <- sample_n(may20044r, 15) 
may20044s <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "4S")
may20044s <- sample_n(may20044s, 29) 
may20044t <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "4T")
may20044t <- sample_n(may20044t, 40) 
may20044vn <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "4Vn")
may20044vn <- sample_n(may20044vn, 8) 
may20044vs <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "4Vs")
may20044vs <- sample_n(may20044vs, 31) 
may20044w <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "4W")
may20044w <- sample_n(may20044w, 32) 
may20044x <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "4X")
may20044x <- sample_n(may20044x, 30) 
may20045y <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "5Y")
may20045y <- sample_n(may20045y, 12) 
may20045ze <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "5Ze")
may20045ze <- sample_n(may20045ze, 4) 
may2004hud <- subset(backno1, year == "2004" & month == "5" & nafo_zone == "HudsonStrait")
may2004hud <- sample_n(may2004hud, 31) 
may2004 <- rbind(may20040a, may20040b, may20041B, may20041C, may20041d, may20041E, may20041F,  may20042G, may20042H, may20042J, may20043k, may20043L, may20043M, may20043N, may20043o, may20043pn, may20043ps, may20044r, may20044s, may20044t, may20044vn, may20044vs, may20044w, may20044x, may20045y, may20045ze, may2004hud)
```

```{r may 2005}
may20050a <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "0A")
may20050a <- sample_n(may20050a, 15) 
may20050b <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "0B")
may20050b <- sample_n(may20050b, 61) 
may20051B <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "1B")
may20051B <- sample_n(may20051B, 1)
may20051C <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "1C")
may20051C <- sample_n(may20051C, 5) 
may20051d <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "1D")
may20051d <- sample_n(may20051d, 10)
may20051E <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "1E")
may20051E <- sample_n(may20051E, 13)
may20051F <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "1F")
may20051F <- sample_n(may20051F, 25)
may20052G <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "2G")
may20052G <- sample_n(may20052G, 44) 
may20052H <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "2H")
may20052H <- sample_n(may20052H, 46) 
may20052J <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "2J")
may20052J <- sample_n(may20052J, 55) 
may20053k <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "3K")
may20053k <- sample_n(may20053k, 57) 
may20053L <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "3L")
may20053L <- sample_n(may20053L, 48)
may20053M <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "3M")
may20053M <- sample_n(may20053M, 3)
may20053N <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "3N")
may20053N <- sample_n(may20053N, 17) 
may20053o <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "3O")
may20053o <- sample_n(may20053o, 25) 
may20053pn <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "3Pn")
may20053pn <- sample_n(may20053pn, 2) 
may20053ps <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "3Ps")
may20053ps <- sample_n(may20053ps, 26) 
may20054r <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "4R")
may20054r <- sample_n(may20054r, 12) 
may20054s <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "4S")
may20054s <- sample_n(may20054s, 24) 
may20054t <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "4T")
may20054t <- sample_n(may20054t, 33) 
may20054vn <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "4Vn")
may20054vn <- sample_n(may20054vn, 7) 
may20054vs <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "4Vs")
may20054vs <- sample_n(may20054vs, 26) 
may20054w <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "4W")
may20054w <- sample_n(may20054w, 26) 
may20054x <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "4X")
may20054x <- sample_n(may20054x, 24) 
may20055y <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "5Y")
may20055y <- sample_n(may20055y, 10) 
may20055ze <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "5Ze")
may20055ze <- sample_n(may20055ze, 3) 
may2005hud <- subset(backno1, year == "2005" & month == "5" & nafo_zone == "HudsonStrait")
may2005hud <- sample_n(may2005hud, 26) 
may2005 <- rbind(may20050a, may20050b, may20051B, may20051C, may20051d, may20051E, may20051F,  may20052G, may20052H, may20052J, may20053k, may20053L, may20053M, may20053N, may20053o, may20053pn, may20053ps, may20054r, may20054s, may20054t, may20054vn, may20054vs, may20054w, may20054x, may20055y, may20055ze, may2005hud)
```

```{r may 2006}
may20060a <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "0A")
may20060a <- sample_n(may20060a, 1) 
may20060b <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "0B")
may20060b <- sample_n(may20060b, 5) 
may20061B <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "1B")
may20061B <- sample_n(may20061B, 0)
may20061C <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "1C")
may20061C <- sample_n(may20061C, 0) 
may20061d <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "1D")
may20061d <- sample_n(may20061d, 1)
may20061E <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "1E")
may20061E <- sample_n(may20061E, 1)
may20061F <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "1F")
may20061F <- sample_n(may20061F, 2)
may20062G <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "2G")
may20062G <- sample_n(may20062G, 3) 
may20062H <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "2H")
may20062H <- sample_n(may20062H,4) 
may20062J <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "2J")
may20062J <- sample_n(may20062J, 4) 
may20063k <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "3K")
may20063k <- sample_n(may20063k, 4) 
may20063L <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "3L")
may20063L <- sample_n(may20063L, 4)
may20063M <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "3M")
may20063M <- sample_n(may20063M, 0)
may20063N <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "3N")
may20063N <- sample_n(may20063N, 1) 
may20063o <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "3O")
may20063o <- sample_n(may20063o, 2) 
may20063pn <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "3Pn")
may20063pn <- sample_n(may20063pn, 0) 
may20063ps <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "3Ps")
may20063ps <- sample_n(may20063ps, 2) 
may20064r <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "4R")
may20064r <- sample_n(may20064r, 1) 
may20064s <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "4S")
may20064s <- sample_n(may20064s, 2) 
may20064t <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "4T")
may20064t <- sample_n(may20064t, 2) 
may20064vn <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "4Vn")
may20064vn <- sample_n(may20064vn, 1) 
may20064vs <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "4Vs")
may20064vs <- sample_n(may20064vs, 2) 
may20064w <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "4W")
may20064w <- sample_n(may20064w, 2) 
may20064x <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "4X")
may20064x <- sample_n(may20064x, 2) 
may20065y <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "5Y")
may20065y <- sample_n(may20065y, 1) 
may20065ze <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "5Ze")
may20065ze <- sample_n(may20065ze, 0) 
may2006hud <- subset(backno1, year == "2006" & month == "5" & nafo_zone == "HudsonStrait")
may2006hud <- sample_n(may2006hud, 2) 
may2006 <- rbind(may20060a, may20060b, may20061B, may20061C, may20061d, may20061E, may20061F,  may20062G, may20062H, may20062J, may20063k, may20063L, may20063M, may20063N, may20063o, may20063pn, may20063ps, may20064r, may20064s, may20064t, may20064vn, may20064vs, may20064w, may20064x, may20065y, may20065ze, may2006hud)
```


```{r may 2007}
may20070a <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "0A")
may20070a <- sample_n(may20070a, 12) 
may20070b <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "0B")
may20070b <- sample_n(may20070b, 49) 
may20071B <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "1B")
may20071B <- sample_n(may20071B, 1)
may20071C <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "1C")
may20071C <- sample_n(may20071C, 4) 
may20071d <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "1D")
may20071d <- sample_n(may20071d, 8)
may20071E <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "1E")
may20071E <- sample_n(may20071E, 10)
may20071F <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "1F")
may20071F <- sample_n(may20071F, 19)
may20072G <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "2G")
may20072G <- sample_n(may20072G, 35) 
may20072H <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "2H")
may20072H <- sample_n(may20072H, 37) 
may20072J <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "2J")
may20072J <- sample_n(may20072J, 43) 
may20073k <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "3K")
may20073k <- sample_n(may20073k, 45) 
may20073L <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "3L")
may20073L <- sample_n(may20073L, 38)
may20073M <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "3M")
may20073M <- sample_n(may20073M, 2)
may20073N <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "3N")
may20073N <- sample_n(may20073N, 14) 
may20073o <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "3O")
may20073o <- sample_n(may20073o, 20) 
may20073pn <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "3Pn")
may20073pn <- sample_n(may20073pn, 2) 
may20073ps <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "3Ps")
may20073ps <- sample_n(may20073ps, 20) 
may20074r <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "4R")
may20074r <- sample_n(may20074r, 10) 
may20074s <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "4S")
may20074s <- sample_n(may20074s, 19) 
may20074t <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "4T")
may20074t <- sample_n(may20074t, 26) 
may20074vn <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "4Vn")
may20074vn <- sample_n(may20074vn, 5) 
may20074vs <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "4Vs")
may20074vs <- sample_n(may20074vs, 20) 
may20074w <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "4W")
may20074w <- sample_n(may20074w, 20) 
may20074x <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "4X")
may20074x <- sample_n(may20074x, 19) 
may20075y <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "5Y")
may20075y <- sample_n(may20075y, 8) 
may20075ze <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "5Ze")
may20075ze <- sample_n(may20075ze, 2) 
may2007hud <- subset(backno1, year == "2007" & month == "5" & nafo_zone == "HudsonStrait")
may2007hud <- sample_n(may2007hud, 20) 
may2007 <- rbind(may20070a, may20070b, may20071B, may20071C, may20071d, may20071E, may20071F,  may20072G, may20072H, may20072J, may20073k, may20073L, may20073M, may20073N, may20073o, may20073pn, may20073ps, may20074r, may20074s, may20074t, may20074vn, may20074vs, may20074w, may20074x, may20075y, may20075ze, may2007hud)
```


```{r may 2008}
may20080a <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "0A")
may20080a <- sample_n(may20080a, 17) 
may20080b <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "0B")
may20080b <- sample_n(may20080b, 72) 
may20081B <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "1B")
may20081B <- sample_n(may20081B, 1)
may20081C <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "1C")
may20081C <- sample_n(may20081C, 6) 
may20081d <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "1D")
may20081d <- sample_n(may20081d, 11)
may20081E <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "1E")
may20081E <- sample_n(may20081E, 15)
may20081F <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "1F")
may20081F <- sample_n(may20081F, 28)
may20082G <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "2G")
may20082G <- sample_n(may20082G, 50) 
may20082H <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "2H")
may20082H <- sample_n(may20082H, 53) 
may20082J <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "2J")
may20082J <- sample_n(may20082J, 63) 
may20083k <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "3K")
may20083k <- sample_n(may20083k, 65) 
may20083L <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "3L")
may20083L <- sample_n(may20083L, 55)
may20083M <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "3M")
may20083M <- sample_n(may20083M, 3)
may20083N <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "3N")
may20083N <- sample_n(may20083N, 20) 
may20083o <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "3O")
may20083o <- sample_n(may20083o, 29) 
may20083pn <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "3Pn")
may20083pn <- sample_n(may20083pn, 3) 
may20083ps <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "3Ps")
may20083ps <- sample_n(may20083ps, 30) 
may20084r <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "4R")
may20084r <- sample_n(may20084r, 14) 
may20084s <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "4S")
may20084s <- sample_n(may20084s, 27) 
may20084t <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "4T")
may20084t <- sample_n(may20084t, 38) 
may20084vn <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "4Vn")
may20084vn <- sample_n(may20084vn, 8) 
may20084vs <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "4Vs")
may20084vs <- sample_n(may20084vs, 30) 
may20084w <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "4W")
may20084w <- sample_n(may20084w, 30) 
may20084x <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "4X")
may20084x <- sample_n(may20084x, 28) 
may20085y <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "5Y")
may20085y <- sample_n(may20085y, 11) 
may20085ze <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "5Ze")
may20085ze <- sample_n(may20085ze, 4) 
may2008hud <- subset(backno1, year == "2008" & month == "5" & nafo_zone == "HudsonStrait")
may2008hud <- sample_n(may2008hud, 30) 
may2008 <- rbind(may20080a, may20080b, may20081B, may20081C, may20081d, may20081E, may20081F,  may20082G, may20082H, may20082J, may20083k, may20083L, may20083M, may20083N, may20083o, may20083pn, may20083ps, may20084r, may20084s, may20084t, may20084vn, may20084vs, may20084w, may20084x, may20085y, may20085ze, may2008hud)
```

```{r may 2009}
may20090a <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "0A")
may20090a <- sample_n(may20090a, 33) 
may20090b <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "0B")
may20090b <- sample_n(may20090b, 136) 
may20091B <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "1B")
may20091B <- sample_n(may20091B, 2)
may20091C <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "1C")
may20091C <- sample_n(may20091C, 12) 
may20091d <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "1D")
may20091d <- sample_n(may20091d, 22)
may20091E <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "1E")
may20091E <- sample_n(may20091E, 29)
may20091F <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "1F")
may20091F <- sample_n(may20091F, 54)
may20092G <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "2G")
may20092G <- sample_n(may20092G, 95) 
may20092H <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "2H")
may20092H <- sample_n(may20092H, 101) 
may20092J <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "2J")
may20092J <- sample_n(may20092J, 119) 
may20093k <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "3K")
may20093k <- sample_n(may20093k, 124) 
may20093L <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "3L")
may20093L <- sample_n(may20093L, 104)
may20093M <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "3M")
may20093M <- sample_n(may20093M, 7)
may20093N <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "3N")
may20093N <- sample_n(may20093N, 38) 
may20093o <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "3O")
may20093o <- sample_n(may20093o, 54) 
may20093pn <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "3Pn")
may20093pn <- sample_n(may20093pn, 5) 
may20093ps <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "3Ps")
may20093ps <- sample_n(may20093ps, 56) 
may20094r <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "4R")
may20094r <- sample_n(may20094r, 27) 
may20094s <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "4S")
may20094s <- sample_n(may20094s, 51) 
may20094t <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "4T")
may20094t <- sample_n(may20094t, 71) 
may20094vn <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "4Vn")
may20094vn <- sample_n(may20094vn, 15) 
may20094vs <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "4Vs")
may20094vs <- sample_n(may20094vs, 56) 
may20094w <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "4W")
may20094w <- sample_n(may20094w, 57) 
may20094x <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "4X")
may20094x <- sample_n(may20094x, 53) 
may20095y <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "5Y")
may20095y <- sample_n(may20095y, 21) 
may20095ze <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "5Ze")
may20095ze <- sample_n(may20095ze, 7) 
may2009hud <- subset(backno1, year == "2009" & month == "5" & nafo_zone == "HudsonStrait")
may2009hud <- sample_n(may2009hud, 56) 
may2009 <- rbind(may20090a, may20090b, may20091B, may20091C, may20091d, may20091E, may20091F,  may20092G, may20092H, may20092J, may20093k, may20093L, may20093M, may20093N, may20093o, may20093pn, may20093ps, may20094r, may20094s, may20094t, may20094vn, may20094vs, may20094w, may20094x, may20095y, may20095ze, may2009hud)
```

```{r may 2010}
may20100a <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "0A")
may20100a <- sample_n(may20100a, 16) 
may20100b <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "0B")
may20100b <- sample_n(may20100b, 64) 
may20101B <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "1B")
may20101B <- sample_n(may20101B, 1)
may20101C <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "1C")
may20101C <- sample_n(may20101C, 6) 
may20101d <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "1D")
may20101d <- sample_n(may20101d, 10)
may20101E <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "1E")
may20101E <- sample_n(may20101E, 14)
may20101F <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "1F")
may20101F <- sample_n(may20101F, 25)
may20102G <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "2G")
may20102G <- sample_n(may20102G, 45) 
may20102H <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "2H")
may20102H <- sample_n(may20102H, 48) 
may20102J <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "2J")
may20102J <- sample_n(may20102J, 56) 
may20103k <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "3K")
may20103k <- sample_n(may20103k, 58) 
may20103L <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "3L")
may20103L <- sample_n(may20103L, 49)
may20103M <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "3M")
may20103M <- sample_n(may20103M, 3)
may20103N <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "3N")
may20103N <- sample_n(may20103N, 18) 
may20103o <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "3O")
may20103o <- sample_n(may20103o, 26) 
may20103pn <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "3Pn")
may20103pn <- sample_n(may20103pn, 2) 
may20103ps <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "3Ps")
may20103ps <- sample_n(may20103ps, 27) 
may20104r <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "4R")
may20104r <- sample_n(may20104r, 13) 
may20104s <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "4S")
may20104s <- sample_n(may20104s, 24) 
may20104t <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "4T")
may20104t <- sample_n(may20104t, 34) 
may20104vn <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "4Vn")
may20104vn <- sample_n(may20104vn, 7) 
may20104vs <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "4Vs")
may20104vs <- sample_n(may20104vs, 26) 
may20104w <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "4W")
may20104w <- sample_n(may20104w, 27) 
may20104x <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "4X")
may20104x <- sample_n(may20104x, 25) 
may20105y <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "5Y")
may20105y <- sample_n(may20105y, 10) 
may20105ze <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "5Ze")
may20105ze <- sample_n(may20105ze, 3) 
may2010hud <- subset(backno1, year == "2010" & month == "5" & nafo_zone == "HudsonStrait")
may2010hud <- sample_n(may2010hud, 26) 
may2010 <- rbind(may20100a, may20100b, may20101B, may20101C, may20101d, may20101E, may20101F,  may20102G, may20102H, may20102J, may20103k, may20103L, may20103M, may20103N, may20103o, may20103pn, may20103ps, may20104r, may20104s, may20104t, may20104vn, may20104vs, may20104w, may20104x, may20105y, may20105ze, may2010hud)
```

```{r may 2011}
may20110a <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "0A")
may20110a <- sample_n(may20110a, 21) 
may20110b <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "0B")
may20110b <- sample_n(may20110b, 85) 
may20111B <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "1B")
may20111B <- sample_n(may20111B, 2)
may20111C <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "1C")
may20111C <- sample_n(may20111C, 7) 
may20111d <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "1D")
may20111d <- sample_n(may20111d, 14)
may20111E <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "1E")
may20111E <- sample_n(may20111E, 18)
may20111F <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "1F")
may20111F <- sample_n(may20111F, 33)
may20112G <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "2G")
may20112G <- sample_n(may20112G, 60) 
may20112H <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "2H")
may20112H <- sample_n(may20112H, 63) 
may20112J <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "2J")
may20112J <- sample_n(may20112J, 75) 
may20113k <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "3K")
may20113k <- sample_n(may20113k, 77) 
may20113L <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "3L")
may20113L <- sample_n(may20113L, 65)
may20113M <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "3M")
may20113M <- sample_n(may20113M, 4)
may20113N <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "3N")
may20113N <- sample_n(may20113N, 24) 
may20113o <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "3O")
may20113o <- sample_n(may20113o, 34) 
may20113pn <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "3Pn")
may20113pn <- sample_n(may20113pn, 3) 
may20113ps <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "3Ps")
may20113ps <- sample_n(may20113ps, 35) 
may20114r <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "4R")
may20114r <- sample_n(may20114r, 17) 
may20114s <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "4S")
may20114s <- sample_n(may20114s, 32) 
may20114t <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "4T")
may20114t <- sample_n(may20114t, 45) 
may20114vn <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "4Vn")
may20114vn <- sample_n(may20114vn, 9) 
may20114vs <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "4Vs")
may20114vs <- sample_n(may20114vs, 35) 
may20114w <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "4W")
may20114w <- sample_n(may20114w, 35) 
may20114x <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "4X")
may20114x <- sample_n(may20114x, 33) 
may20115y <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "5Y")
may20115y <- sample_n(may20115y, 13) 
may20115ze <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "5Ze")
may20115ze <- sample_n(may20115ze, 4) 
may2011hud <- subset(backno1, year == "2011" & month == "5" & nafo_zone == "HudsonStrait")
may2011hud <- sample_n(may2011hud, 35) 
may2011 <- rbind(may20110a, may20110b, may20111B, may20111C, may20111d, may20111E, may20111F,  may20112G, may20112H, may20112J, may20113k, may20113L, may20113M, may20113N, may20113o, may20113pn, may20113ps, may20114r, may20114s, may20114t, may20114vn, may20114vs, may20114w, may20114x, may20115y, may20115ze, may2011hud)
```

```{r merge all may}
may <- rbind(may1998, may1999, may2000, may2001, may2002, may2003, may2004, may2005, may2006, may2007, may2008, may2009, may2010, may2011)
```






join everything together
```{r}
back10k <- rbind(jan, feb, mar, apr, may, jun, jul, aug, sep, oct, nov, dec)
write.csv(back10k, "../output/bio/background_10k.csv", row.names = FALSE)
```

merge with presence

now merge the two datasets
```{r merge pres and back10k}
presab10k <- rbind.fill(presencemerging, back10k)
write.csv(presab10k, "../output/bio/presab_10k.csv", row.names = FALSE)
nrow(presab10k) #130300
```

