---
title: "netcdf_to_brick_projected"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
Mass netCDF to raster brick, and then reprojection...

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. The code herein is unlikely to be elegant and there is likely more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.

```{r pre-install packages, message=FALSE}
packages <- c("ncdf4")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

Ok first get a list of netcdfs in the folder...
```{r}
netcdf_list <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = TRUE)
netcdf_list
```

ok now create a raster brick for each netcdf (with the same file name),

```{r}
no_netcdf <- length(netcdf_list)
netcdf_name <- list.files("../data/env/netcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
library (ncdf4)
for (i in 1:no_netcdf) {
  print(netcdf_name[i]) #this just prings the name of the netCDF R is working on
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  assign(netcdf_name[i], temp_brick) #this asigns the netCDF name to the raster brick
}

```



