---
title: "parallel for the env. data grab"
author: "Samantha Andrews"
output: 
  html_notebook: 
editor_options: 
  chunk_output_type: inline
---

# Overview
Attempt to run my inefficinet giant loop for grabbing env. data to the background points in parallel (need to run again AND maybe need to do more points that will currently take 2 months

A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there are probably more efficient ways of running the code.

Built with 'r getRversion()'.

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
The "include=FALSE" supresses the package installation text appearing in the document...
```{r pre-install packages, include=FALSE}
packages <- c("ncdf4", "raster", "foreach", "doParallel") 
source("../src/ipak.R")
ipak(packages)
```

# this test will use a shorter dataset and a smaller number of netcdfs

```{r}
test_back_or <- read.csv("../data/looptest.csv", header = TRUE) #test dataset
xy <- test_back_or[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
test_back <- SpatialPointsDataFrame(coords = xy, data = test_back_or, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.
```


# the inefficent loop

This loop runs through the netcdf files and then looks for which rows in data_aea it should extract the value to point from, and at what depth (netcDF layer)

```{r}
strt <- Sys.time() #get the start time
netcdf_list <- list.files("../data/testloopnetcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/testloopnetcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("../output/env/aea.tif") 
for (i in 1:no_netcdf) {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    for (j in 1:nrow(test_back)) {  
      de <- test_back$depthlayerno[[j]]  # a variable for the observation depth layer
      yr <- (test_back$year[j])  # a variable for the observation year
      mth <- (test_back$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
              test_back$temp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$temp_depth[j] <- NA
              } else  
                test_back$temp_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
              test_back$salinity_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$salinity_depth[j] <- NA
              } else  
                test_back$salinity_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
              test_back$chl_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$chl_depth[j] <- NA
              } else  
                test_back$chl_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
              test_back$o2_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$o2_depth[j] <- NA
              } else  
                test_back$o2_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
              test_back$mlp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
              test_back$ssh_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
            
          }
     
    }
}
write.csv(test_back, "../data/testloopoutput.csv", row.names = FALSE)
test_back_df <- as.data.frame(test_back)
head(test_back_df)
print(Sys.time()-strt) #time it took to run
```


# edit test 1


```{r}
test_back_or <- read.csv("../data/looptest.csv", header = TRUE) #test dataset
xy <- test_back_or[ ,c("longitude_meters","latitude_meters")] # This is to tell R where the coordinates are. Note that the column order needs to be longitude, latitude
test_back <- SpatialPointsDataFrame(coords = xy, data = test_back_or, proj4string = CRS("+proj=aea +lat_1=50 +lat_2=70 +lat_0=40 +lon_0=-60 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")) # The CRS is used here is for the albers equal area projection.

ncores <- detectCores()
cl <- (ncores)-1
registerDoParallel(cl)

strt <- Sys.time() #get the start time

netcdf_list <- list.files("../data/testloopnetcdf", pattern = '*.nc', full.names = TRUE) #true means the full path is included
no_netcdf <- length(netcdf_list) #for the loop - need to know how many files to cycle through
netcdf_name <- list.files("../data/testloopnetcdf", pattern = '*.nc', full.names = FALSE) #false means the path is not included
aea <- raster("../output/env/aea.tif") 
foreach ((i = 1:no_netcdf) , .packages = "raster") %dopar% {  
  print(netcdf_name[i]) #this just prints the name of the netCDF R is working one
  brkyr <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 1)) # extracting the first part of the netcdf filename (which is the year)
  brkmth <- as.integer(sapply(strsplit(netcdf_name[i], "_"), "[[", 2)) # extracting the second part of the netcdf filename (which is the month)
  brkvar <- (sapply(strsplit(netcdf_name[i], "_"), "[[", 3)) # extracting the third part of the netcdf (inc.nc)
  temp_brick <- brick(netcdf_list[i], lvar = 4)
  temp_brick <- projectRaster(temp_brick, aea) 
    foreach (j = 1:nrow(test_back)) %dopar% {  
      de <- test_back$depthlayerno[[j]]  # a variable for the observation depth layer
      yr <- (test_back$year[j])  # a variable for the observation year
      mth <- (test_back$month[j])  # a variable for the observation month
          if (brkyr == yr & brkmth == mth & brkvar == "temp.nc"){
              test_back$temp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$temp_depth[j] <- NA
              } else  
                test_back$temp_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "salinity.nc") {
              test_back$salinity_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$salinity_depth[j] <- NA
              } else  
                test_back$salinity_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "chl.nc") {
              test_back$chl_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$chl_depth[j] <- NA
              } else  
                test_back$chl_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "o2.nc") {
              test_back$o2_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
              if (is.na(de)){
                test_back$o2_depth[j] <- NA
              } else  
                test_back$o2_depth[j] <- extract(x=temp_brick[[de]], y = test_back[j, ]) 
          } else if (brkyr == yr & brkmth == mth & brkvar == "mlp.nc") {
              test_back$mlp_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ])
          } else if (brkyr == yr & brkmth == mth & brkvar == "ssh.nc") {
              test_back$ssh_surface[j] <- extract(x=temp_brick[[1]], y = test_back[j, ]) 
            
          }
     
    }
}
write.csv(test_back, "../data/testloopoutput.csv", row.names = FALSE)
test_back_df <- as.data.frame(test_back)
head(test_back_df)
stopCluster(cl)
print(Sys.time()-strt) #time it took to run
```

